### 2020-03-30 Rmd/Ideas.Rmd
Metadata almost ready, lacking info of 3 networks

---
title: "Ideas"
output: html_notebook
---

#How to proceed

- Select main floral traits

1) Flower size

2) Fower shape

3) Compatibility system

4) Style length/Ovule number

5) Selfing level

- Plot taxa/traits Barplot?

Y axis Spp richness? X Axis the grouping variables?

Consider Poll taxa here with more developped plots?

- Hypothesis

 1) Do we have matching between specific taxa and grouping traits? Maybe show this with likelihood of visitation
    1.1 Are selfers/self compatible species less relevant in the network? How can we interpretate that (generalist vs specialization/connectance) We expect less dependency on pollinators...

1.2

 2)

 3) Possibility of testing redundancy with phylogeny

 - How to group Poll taxa.

 What to do with non-insects?