### 2020-04-06 Rmd/Networks_Information.Rmd
Metadata and complementary rmd file explaining the processing of each network

---
output: html_document
fontsize: 12pt
header-includes:
    - \usepackage{setspace}\doublespacing
---

#Networks information

Here I'm going to record the processing history and adittional information, complementary to the metadata of the networks. Also I will justify when we group the networks as metaweb.

__1 carpobrotus_metaweb_Bartomeus_2008__

We have build a metaweb with three different networks that share most of the plant species.

__2_metaweb_opuntia_Bartomeus_2008__

We have build a metaweb with three different networks that share most of the plant species.

__3_bek_2006__

Master student of Jens Olesen that built an entire network for his thesis.

__4_bundgaard_2003__

Master student of Jens Olesen that built a network for his thesis, sampling on the sides of the road. Quite close to Bek's 2006 network (few hundred meters away).

__5_metaweb_chacoff_2011__

Metaweb from journal without modifications.

__6_metaweb_dicks_2002__

We have combined the two networks provided in this journal article in a unique metaweb. The meadows share a great number of species and in the article they are described as similar systems. From google maps, they seems to be 60-80 kms away from each other.

__7_metaweb_dupont_2009__

There are three networks available with three different sites (obtained from online databases). However, with quantitative data (visitation) just one from the article, seems to be an error with the supplementary material. I have no been able to get the three of them.

__8_elberling_1999__

One unique network with quantitative data. I have passed it manually from the original paper due to the online databases have mistakes in the data and binary.

__9_dupont_2009_zackenberg__

Unique network sampled during one season.

__10_metaweb_fang_huang_2012__

I have combined the data obtained from the journal article in a unique metaweb. There three networks from 3 different seasons.

__11_metaweb_inouye_1990__

Metaweb with multiple seasons. No modifications made.

__12_inouye_1988__

One unique network without modifications.

__13_metaweb_kaiser_bunbury_2009__

One unique network sampled during one season.

__14_metaweb_kaiser_bunbury_2014__

I have merged the 48 sites sampled in a heterogeneous landscape and created a metaweb. Just 6 different sites.

__15_metaweb_kato_2000__

Metaweb obtained from web of life. I have not done modifications to the dataset. Quantitative data.

__16_kevan_1970__

One unique network from one season. I have not modified the data. Obtained from web of life.

__17_lundgren_2005__

One unique network from one season. I have not modified the data.

__18_olesen_2002_mauritius__

One unique network, I have passed manually in order to have the visitation data.

__19_mcmullen_1993__

Metaweb from all the islands. From web of life.

__20_metaweb_arthurs_pass_primack_1983__

Netwrok from web of life, binary data. I have not modified it.

__21_metaweb_cass_primack_1983__

Netwrok from web of life, binary data. I have not modified it.

__22_metaweb_craigieburn_primack_1983__

Netwrok from web of life, binary data. I have not modified it.

__23_ramirez_1989__

Data from web of life. I have not modified the data.

__24_metaweb_ramirez_1992__

Metaweb from web of life without modifications.

__25_metaweb_robertson_1929__

Metaweb from web of life without modifications.

__26_small_1976__

Unique network with quantitative data without modifications.

__27_chaco_souza_2018__

Unique network with quantitative data without modifications.

__28_metaweb_traveset_2013__

Metaweb with networks from different islands from journal.

__29_metaweb_bartomeus_unpublished_data_2015__

Metaweb with networks from different sites. I made a subset of this data by transects.

__30_olesen_2002_azores__

Unique network without modifications. I have passed the data manually from source.