### 2020-08-07 Rmd/Metadata_table.tex
brms model

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Untitled},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Untitled}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{landscape}

\begingroup\fontsize{7}{9}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Id & Longitude & Latitude & Country & Year & Networks & Plant spp & Pollinator spp & Network size & Sampling method & Sampling & Data type\\
\midrule
\endfirsthead
\multicolumn{12}{@{}l}{\textit{(continued)}}\\
\toprule
Id & Longitude & Latitude & Country & Year & Networks & Plant spp & Pollinator spp & Network size & Sampling method & Sampling & Data type\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\rowcolor{gray!6}  Bartomeus 2008 & 24 & 18.0 & Spain & 2005 & 3 & 18 & 37 & 666 & Plots with transects & Phytocentric & Quantitative\\
\addlinespace
Fang 2012 & 29 & 12.0 & China & 2008-2010 & 3 & 130 & 247 & 32110 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Inouye 1990 & 18 & 14.0 & Japan & 1984-1987 & 4 & 114 & 883 & 100662 & Transects & Phytocentric & Quantitative\\
\addlinespace
Inouye 1988 & 19 & 7.0 & Australia & 1983-1984 & 1 & 40 & 85 & 3400 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Kaiser-Bunbury 2009 & 27 & 4.0 & Republic of Mauritius & 2003-2004 & 2 & 96 & 184 & 17664 & Plots & Phytocentric & Quantitative\\
\addlinespace
Kaiser-Bunbury 2014 & 25 & 8.0 & Republic of Seychelles & 2007-2008 & 6 & 37 & 341 & 12617 & Transects & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Kato 2000 & 17 & 13.0 & Japan & 1996-1999 & 16 & 110 & 609 & 66990 & Transects & Phytocentric & Quantitative\\
\addlinespace
Kevan 1970 & 9 & 28.0 & Canada & 1967 & 1 & 20 & 91 & 1820 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
\rowcolor{gray!6}  Lundgren 2005 & 3 & 25.0 & Greenland & 2002 & 1 & 17 & 26 & 442 & Randow census walks & Phytocentric & Quantitative\\
\addlinespace
Olesen 2002 1 & 26 & 3.0 & Republic of Mauritius & 1998-1999 & 1 & 17 & 26 & 442 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Mcmullen 1993 & 12 & 1.0 & Ecuador & NA & All islands & 105 & 54 & 5670 & - & Phytocentric & Qualitative\\
\addlinespace
Bartomeus 2008 & 24 & 18.0 & Spain & 2005 & 3 & 13 & 37 & 481 & Plots with transects & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Primack 1983 1 & 20 & 9.0 & New Zealand & 1976-1978 & 1 & 18 & 60 & 1080 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
Primack 1983 2 & 22 & 10.0 & New Zealand & 1976-1978 & 1 & 41 & 139 & 5699 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
\rowcolor{gray!6}  Primack 1983 3 & 21 & 11.0 & New Zealand & 1976-1978 & 1 & 49 & 118 & 5782 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
Ramirez 1989 & 6 & 20.0 & Venezuela & NA & 1 & 48 & 49 & 2352 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
\rowcolor{gray!6}  Ramirez 1992 & 7 & 27.0 & Venezuela & 1983,1984,1989 & 1 & 28 & 53 & 1484 & Randow census walks & Phytocentric & Qualitative\\
\addlinespace
Robertson 1929 & 11 & 16.0 & United States & 1997-1899 & NA & 456 & 1044 & 476064 & - & Phytocentric & Qualitative\\
\addlinespace
\rowcolor{gray!6}  Small 1976 & 10 & 19.0 & Canada & 1973 & 1 & 13 & 34 & 442 & 10h per spp & Phytocentric & Quantitative\\
\addlinespace
Souza 2018 & 4 & 5.0 & Brazil & 2008-2009 & 1 & 62 & 89 & 5518 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Traveset 2013 & 13 & 2.0 & Ecuador & 2010-2011 & 1 & 60 & 220 & 13200 & Randow census walks & Phytocentric & Quantitative\\
\addlinespace
Bartomeus 2015 unp. & 5 & 15.0 & Spain & 2015 & 16 & 57 & 277 & 15789 & Transects & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Bek 2006 & 15 & 22.0 & Denmark & 2003 & 1 & 37 & 225 & 8325 & Plots & Phytocentric & Qualitative\\
\addlinespace
Olesen 2002 2 & 2 & 17.0 & Azores & 2000 & 1 & 10 & 12 & 120 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Bundgaard 2003 & 16 & 22.0 & Denmark & 2003 & 1 & 16 & 44 & 704 & Plots & Phytocentric & Qualitative\\
\addlinespace
Chacoff 2011 & 8 & 6.0 & Argentina & 2006-2009 & 4 & 59 & 196 & 11564 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Dicks 2002 & 14 & 21.0 & England & 2001? & 2 & 23 & 80 & 1840 & Plots & Phytocentric & Quantitative\\
\addlinespace
Dupont 2009 & 28 & 56.1 & Denmark & 2005 & 2 & 31 & 329 & 10199 & Plots & Phytocentric & Quantitative\\
\addlinespace
\rowcolor{gray!6}  Elberling 1999 & 23 & 24.0 & Sweden & 1994 & 1 & 24 & 118 & 2832 & Plots with transects & Phytocentric & Quantitative\\
\addlinespace
Dupont 2009 2 & 1 & 26.0 & Greenland & 1996-1997 & 1 & 31 & 76 & 2356 & Random census walks & Phytocentric & Qualitative\\*
\end{longtabu}
\endgroup{}
\end{landscape}

\end{document}