### 2020-09-09 Rmd/brms_metric_selfing.Rmd
preparing output visits inteher/selfing level

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

# Model 1 
# $$Visits / SelfingLevel + (1|net_id) + (1|gr(phylo, cov = A))$$

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/RData/brms_m1_visits_self_all.RData")
load("Data/RData/brms_data_m1_visits_self_all.RData")

c_e <- conditional_effects(m1)
p1 <- plot(c_e, points=T,plot = FALSE)[[1]]

#Model fit
pp_check(m1) + xlim(-100,2000)+ylim(0,0.02)
#Estimates, se and ci
d <- p1[[1]][,c(1,7:10)]
kable(d, "latex")

```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/RData/brms_m1_visits_self_all.RData")
load("Data/RData/brms_data_m1_visits_self_all.RData")

c_e <- conditional_effects(m1)
p1 <- plot(c_e, points=T,plot = FALSE)[[1]]

#Prepare example with selfing level
all_df_2 <- all_df_2 %>%
  mutate(autonomous_selfing_level = fct_relevel(autonomous_selfing_level, levels=c("high", "medium", "low", "none")))
all_df_2$autonomous_selfing_level <- as.factor(all_df_2$autonomous_selfing_level)

#Log
ggplot(data=p1[[1]], aes(x = autonomous_selfing_level, y = log(visits),color = ordered(autonomous_selfing_level))) +
  geom_point(data = all_df_2,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p1[[1]],mapping=aes(x=autonomous_selfing_level, ymin=log(lower__), ymax=log(upper__)), width=.1, color="black")+
  geom_point(data=p1[[1]], mapping=aes(x=autonomous_selfing_level, y=log(estimate__)), color="black") + ylab("Visits") + xlab("Selfing level")+
  theme(legend.position = "none")

```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/RData/brms_m1_visits_self_all.RData")
load("Data/RData/brms_data_m1_visits_self_all.RData")

c_e <- conditional_effects(m1)
p1 <- plot(c_e, points=T,plot = FALSE)[[1]]

#ylim500
ggplot(data=p1[[1]], aes(x = autonomous_selfing_level, y = visits,color = ordered(autonomous_selfing_level))) +
  geom_point(data = all_df_2,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p1[[1]],mapping=aes(x=autonomous_selfing_level, ymin=lower__, ymax=upper__), width=.1, color="black")+
  geom_point(data=p1[[1]], mapping=aes(x=autonomous_selfing_level, y=estimate__), color="black") + ylab("Visits") + xlab("Selfing level")+
  theme(legend.position = "none")+ylim(0,500)

```