### 2020-09-17 Visits_Traits.Rmd
Preparing summary of what I have found

---
header-includes:
    - \usepackage{setspace}\doublespacing
output:
  pdf_document: default
  html_document: default
---

\setlength{\abovedisplayskip}{25pt}
\setlength{\belowdisplayskip}{25pt}
\setlength{\abovedisplayshortskip}{25pt}
\setlength{\belowdisplayshortskip}{25pt}

```{r, echo = F}
library(knitr)
if(is_latex_output()) {
  plot_default <- knit_hooks$get("plot")
  knit_hooks$set(plot = function(x, options) { 
    x <- c(plot_default(x, options), "\\vspace{25pt}")
  })
}
```

#COMPATIBILITY

##VISITS~COMPATIBILITY

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Compatibility/brms_m1_visits_compatibility_negative_binomial.RData")
load("Data/Brms/Compatibility/brms_data_compatibility.RData")
c_e <- conditional_effects(m1)
p1 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT

ggplot(data=p1[[1]], aes(x = compatibility, y = log(visits),color = ordered(compatibility))) +
  geom_point(data = all_df_3,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p1[[1]],mapping=aes(x=compatibility, ymin=log(lower__), ymax=log(upper__)), width=.1, color="black")+
  geom_point(data=p1[[1]], mapping=aes(x=compatibility, y=log(estimate__)), color="black") + ylab("log(Visits)") + xlab("Compatibility")+
  theme(legend.position = "none")

```

##Z-SCORES~COMPATIBILITY

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Compatibility/brms_m2_z-scores_compatibility_student.RData")
load("Data/Brms/Compatibility/brms_data_compatibility.RData")
c_e <- conditional_effects(m2)
p2 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT

ggplot(data=p2[[1]], aes(x = compatibility, y = z_score_sum,color = ordered(compatibility))) +
  geom_point(data = all_df_3,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p2[[1]],mapping=aes(x=compatibility, ymin=lower__, ymax=upper__), width=.1, color="black")+
  geom_point(data=p2[[1]], mapping=aes(x=compatibility, y=estimate__), color="black") + ylab("Z-score") + xlab("Compatibility")+
  theme(legend.position = "none")

```

\newpage

#AUTONOMOUS SELFING FRUIT SET

##VISITS~SELFING:QUALITATIVE DATA

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Selfing/brms_m1_visits_selfing_negative_binomial.RData")
load("Data/Brms/Selfing/brms_data_m1_selfing.RData")
c_e <- conditional_effects(m1)
p1 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT

ggplot(data=p1[[1]], aes(x = autonomous_selfing_level, y = log(visits),color = ordered(autonomous_selfing_level))) +
  geom_point(data = all_df_2,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p1[[1]],mapping=aes(x=autonomous_selfing_level, ymin=log(lower__), ymax=log(upper__)), width=.1, color="black")+
  geom_point(data=p1[[1]], mapping=aes(x=autonomous_selfing_level, y=log(estimate__)), color="black") + ylab("log(Visits)") + xlab("Autonomous selfing level")+
  theme(legend.position = "none")

```

##VISITS~SELFING:QUANTITATIVE DATA

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Selfing/brms_m3_visits_selfing_negative_binomial.RData")
load("Data/Brms/Selfing/brms_data_m3_selfing.RData")
c_e <- conditional_effects(m3)
p3 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT

ggplot(data=p3[[1]], aes(x = autonomous_selfing_level_fruit_set, y = log(visits))) +
geom_point(data = all_df_3,alpha = 1/4) + 
scale_fill_brewer(palette = "Greys") +
scale_color_brewer(palette = "Set2") + theme_bw() + geom_smooth(data = p3[[1]],
aes(y = log(estimate__), ymin = log(lower__), ymax = log(upper__)),stat = "identity", color = "black", alpha = 0.1, size = 1/2)

```

\newpage

#AUTONOMOUS SELFING FRUIT SET

##Z-SCORES~SELFING:QUALITATIVE DATA

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Selfing/brms_m4_z-scores_selfing_student.RData")
load("Data/Brms/Selfing/brms_data_m4_selfing.RData")
c_e <- conditional_effects(m4)
p4 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT
ggplot(data=p4[[1]], aes(x = autonomous_selfing_level, y = z_score_sum,color = ordered(autonomous_selfing_level))) +
  geom_point(data = all_df_2,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() +
  geom_errorbar(data=p4[[1]],mapping=aes(x=autonomous_selfing_level, ymin=(lower__), ymax=(upper__)), width=.1, color="black")+
  geom_point(data=p4[[1]], mapping=aes(x=autonomous_selfing_level, y=(estimate__)), color="black") + ylab("Z-score") + xlab("Selfing level")+theme(legend.position = "none")

```

##Z-SCORES-SELFING:QUALITATIVE DATA

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=7, fig.height=3, fig.align='center'}
library(brms)
library(ggplot2)
library(tidybayes)
library(tidyverse)
library(kableExtra)

setwd("~/R_Projects/Reproductive traits") 
load("Data/Brms/Selfing/brms_m5_z-scores_selfing_student.RData")
load("Data/Brms/Selfing/brms_data_m5_selfing.RData")
c_e <- conditional_effects(m5)
p5 <- plot(c_e, points=T,plot = FALSE)[[1]]

#PLOT OUTPUT

ggplot(data=p5[[1]], aes(x = autonomous_selfing_level_fruit_set, y = z_score_sum)) +
  geom_point(data = all_df_3,alpha = 1/4) + 
  scale_fill_brewer(palette = "Greys") +
  scale_color_brewer(palette = "Set2") + theme_bw() + geom_smooth(data = p5[[1]],
  aes(y = estimate__, ymin = lower__, ymax = upper__),stat = "identity", color = "black", alpha = 0.1, size = 1/2)+
  ylab("Z-scores")+ xlab("Autonomous selfing-fruit set")

```