### 2020-11-12 Rmd/Quantitative_metadata.Rmd
Data

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

\begin{landscape}

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE,}
setwd("~/R_Projects/Reproductive Traits")
library(kableExtra)

Author <- c("Bartomeus", "Dicks", "Dupont", "Elberling", "Fang", "Inouye", "Lundgren", "Olesen", "Small", "Souza", "Kaiser-Bunbury",
            "Bartomeus", "Kaiser-Bunbury", "Kaiser-Bunbury", "Peralta", "Burkle", "Arroyo-Correa", "Bundgaard")

Year <- c("2008", "2002", "2009", "1999", "2008", "1988", "2005", "2002", "1976", "2008", "2017", "2008", "2011", "2010", "2020", "2013", "2019", "2003")

DOI <- c("https://doi.org/10.1007/s00442-007-0946-1", "https://doi.org/10.1046/j.0021-8790.2001.00572.x", "https://doi.org/10.1111/j.1365-2656.2008.01501.x", "https://doi.org/10.1111/j.1365-2656.2008.01501.x", "https://doi.org/10.1111/j.1365-2656.2008.01501.x", "https://doi.org/10.1111/j.1442-9993.1988.tb00968.x",
"https://doi.org/10.1657/1523-0430(2005)037[0514:TDAHCW]2.0.CO;2", 
"https://doi.org/10.1046/j.1472-4642.2002.00148.x", "/13960/t4km08d21", "https://doi.org/10.1111/1365-2745.12978",
"https://doi.org/10.1038/nature21071", "https://github.com/ibartomeus/BeeFunData", "https://doi.org/10.1111/j.1365-2745.2010.01732.x",
"https://doi.org/10.1016/j.ppees.2009.04.001", "https://doi.org/10.1111/ele.13510", "https://doi.org/10.1126/science.1232728", 
"https://doi.org/10.1111/1365-2745.13332", "Unpublished")

Networks_used <- c("6", "2", "2", "1", "1", "1", "1", "2", "1", "1", "8", "9", "6", "2", "4", "1", "3", "1")

Sampling_method <- c("Transects", "Plots", "Plots", "Transects", "Plots", "Plots", "Random census walks", "Plots", "10h per species", "Plots", "Scattered along transects", "Transects", "30min x 4", "Transects", "5min/spp x weekly", "Random census walks", "Transects", "Plots")

m <- data.frame(Author, Year, DOI, Networks_used, Sampling_method)
colnames(m) <- c("Author", "Year", "DOI", "N. of Networks included", "Sampling method")

kable(m, "latex", longtable = T, booktabs = T,linesep = "\\addlinespace") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 7, full_width=T) %>% column_spec(3,width="9cm")
```
\end{landscape}