### 2020-12-14 Scripts/7_1_Visualize_Plant_Functional_Groups/7_plots.Rmd
edits

---
title: "Untitled"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=6}
setwd("~/R_Projects/Reproductive Traits")
library(dplyr)
library(ggplot2)
library(viridis)
library(grid)
library(cowplot)
library(gridExtra)

#PREPARE DATA
#read unscaled trait data in order to visualize better the clusters
d <- read.csv("Data/Csv/all_species_imputed_trait_data.csv")
#read trait data with clusters
#I'll select the column of clusters and include it on the non standardize trait data
hclust_d_5 <- read.csv("Data/Csv/imputed_trait_data_hclust_5_clusters.csv") #5 clusters
#No add the cluster columns (it has the same order)
d$Clusters <- as.factor(hclust_d_5$Clusters)

#select columns of interest
t <- d[c("Breeding_system","IMPUTED_Compatibility","Autonomous_selfing_level",
                        "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Flowers_per_inflorescence",
                        "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean", "STYLE_IMPUTED", "OVULES_IMPUTED", "life_form", "lifespan",
                        "IMPUTED_plant_height_mean_m", "Clusters")]

#GENERATE PLOTS
########################################################################################################################################
#Breeding system
########################################################################################################################################
breeding <- ggplot(t, aes(Clusters, Breeding_system)) +
  geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
  theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Breeding system")+
  geom_hline(yintercept=c(1.5,2.5),color="black",linetype="longdash")
########################################################################################################################################
#Compatibility
########################################################################################################################################
t$IMPUTED_Compatibility <- as.character(t$IMPUTED_Compatibility)
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="monoecious"] <- "Unisexual flowers"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="dioecious"] <- "Unisexual flowers"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_incompatible"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_compatible"] <- "Self compatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="partially_self_compatible"] <- "Partially self compatible"

Compatibility <- ggplot(t, aes(Clusters, IMPUTED_Compatibility)) +
  geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
  theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Compatibility system")+
  geom_hline(yintercept=c(1.5,2.5,3.5),color="black",linetype="longdash")
########################################################################################################################################
#Autonomous_selfing_level
########################################################################################################################################
t$Autonomous_selfing_level <- as.character(t$Autonomous_selfing_level)
t$Autonomous_selfing_level[t$Autonomous_selfing_level=="none"] <- "None"
t$Autonomous_selfing_level[t$Autonomous_selfing_level=="medium"] <- "Medium"
t$Autonomous_selfing_level[t$Autonomous_selfing_level=="low"] <- "Low"
t$Autonomous_selfing_level[t$Autonomous_selfing_level=="high"] <- "High"
t$Autonomous_selfing_level <- factor(t$Autonomous_selfing_level, levels = c("None", "Low", "Medium", "High"))

Selfing <- ggplot(t, aes(Clusters, Autonomous_selfing_level)) +
  geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
  theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Selfing level")+
  geom_hline(yintercept=c(1.5,2.5,3.5),color="black",linetype="longdash")
#######################################################################################################################################
#Flower_morphology
#######################################################################################################################################

t$Flower_morphology <- as.character(t$Flower_morphology)
t$Flower_morphology[t$Flower_morphology=="Funnelform"] <- "Tube"
t$Flower_morphology[t$Flower_morphology=="Spike"] <- "Brush"

Flower_morphology <-  ggplot(t, aes(Clusters, Flower_morphology)) +
  geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
  theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Flower shape")+
  geom_hline(yintercept=c(1.5,2.5,3.5,4.5,5.5),color="black",linetype="longdash")
#######################################################################################################################################
#Flower_symmetry
#######################################################################################################################################
t$Flower_symmetry <- as.character(t$Flower_symmetry)
t$Flower_symmetry[t$Flower_symmetry=="zygomorphic"] <- "Zygomorphic"
t$Flower_symmetry[t$Flower_symmetry=="actinomorphic"] <- "Actinomorphic"

Flower_symmetry <-  ggplot(t, aes(Clusters, Flower_symmetry)) +
   geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
   theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Flower symmetry")+
  geom_hline(yintercept=c(1.5),color="black",linetype="longdash")
#######################################################################################################################################
 #life_form
#######################################################################################################################################
t$life_form <- as.character(t$life_form)
t$life_form[t$life_form=="vine"] <- "Herb"
t$life_form[t$life_form=="herb"] <- "Herb"
t$life_form[t$life_form=="shrub"] <- "Shrub"
t$life_form[t$life_form=="tree"] <- "Tree"

life_form <-  ggplot(t, aes(Clusters,life_form )) +
   geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
   theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x=NULL,y=NULL,subtitle = "Life form")+
  geom_hline(yintercept=c(1.5,2.5),color="black",linetype="longdash")
#######################################################################################################################################
 #lifespan
#######################################################################################################################################
lifespan<- ggplot(t, aes(Clusters,lifespan )) +
   geom_jitter(aes(color = Clusters), size = 0.4)+ scale_color_manual(values =  c("#00AFBB", "#E69F00", "#FC4E07","#000000", "#009E73"))+
   theme_classic()+ theme_classic()+theme(legend.position = "none")+labs(x="Plant Functional groups",y=NULL,subtitle = "Life span")+
  geom_hline(yintercept=c(1.5),color="black",linetype="longdash")

#######################################################################################################################################
 plot_grid(breeding,Compatibility,Selfing,Flower_morphology,Flower_symmetry,life_form,lifespan, nrow = 7,align = 'v')

```