### 2021-03-23 Figures/Figure_Posthoc_Brms_Model.Rmd

---
title: ''
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13}

#LOAD LIBRARIES
library(ape) #for phylogenetic distance
library(dplyr) #data processing
library(rtrees) #for phylogenetic distance
library(DHARMa)
library(brms)
library(cmdstanr)
library(ggplot2)
library(patchwork)
library(tidybayes)
library(emmeans)

theme_ms <- function(base_size=12, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(1.8)),
           axis.text=element_text(size = rel(1.2), color = "black"),
           legend.title=element_text(face="bold"),
           legend.text=element_text(size = rel(1.2)),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(1, 'lines'),
           panel.border=element_rect(color="black",size=1),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank()
     ))
}

setwd("~/Dropbox/PhD/R") #DROPBOX, files too large for github

result <- readRDS("result_Tukey_model.RDS")

ggplot(result, aes(x = .value, y = contrast,  fill=contrast)) +
  stat_eye( point_interval = median_qi,
            .width = c(0.66, 0.95),alpha=0.7) +theme_ms() +theme(legend.position = "none")+geom_vline(xintercept=0,linetype="dashed") + ylab("Pairwise comparisons") + xlab("Marginal mean difference")

```