### 2021-03-23 Figures/BRMS_Metric_PPCA.Rmd
Edits

---
title: ""
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=13, fig.width=17}

library(brms)
library(cmdstanr)
library(cowplot) #panel of plots
library(ggplot2)
library(data.table)

# Theme for publication
theme_ms <- function(base_size=12, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(1.8)),
           axis.text=element_text(size = rel(1.5), color = "black"),
           legend.title=element_text(face="bold"),
           legend.text=element_text(),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(0.4, 'lines'),
           panel.border=element_rect(color="black",size=1),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank()
     ))
}

########################################################################################################################################################################
setwd("~/R_Projects/Reproductive Traits")

#LOAD DATA
data <- read.csv("Data/Csv/metric_analysis_data_3rd_question.csv", row.names = 1) # data with specialization and integer of visits
head(data) #Check data
#Now we obtain the loadings from the global PCA which is likely to define better in the trait space the species
phyl_pca_2 <- readRDS("Data/RData/phyl_pca_forest.rds")
########################################################################################################################################################################
#PROCESS DATA BEFORE MODELLING
pca_loadings_2 <- data.frame(phyl_pca_2$S)
#prepare cols for merging
rownames(pca_loadings_2) <- gsub("_", " ", rownames(pca_loadings_2)) #remove underscores from species names
#set same colnames for merging
pca_loadings_2 <- setDT(pca_loadings_2, keep.rownames = TRUE)[]
#same colname
colnames(pca_loadings_2)[1] <- "Species"
#check data for merge
head(data)
#merge columns
data_analysis2 <- merge(data, pca_loadings_2, by="Species")

setwd("~/Dropbox/PhD/R") #DROPBOX, files too large for github
#LOAD DATA
ce_1 <- readRDS("Visits_PC1.RDS")
ce_2 <- readRDS("Visits_PC2.RDS")
spe_1 <- readRDS("Specialization_PC1.RDS")
spe_2 <- readRDS("Specialization_PC2.RDS")

colnames(ce_1[[1]])[2] <- "Interaction"

Visits_PC1 <- ggplot(ce_1[[1]], aes(x = PC1, y = (estimate__+1))) + geom_point(data = data_analysis2,aes(x = PC1, y = Visits),
       size = 1.5, alpha=0.9) + geom_line(colour="darkblue",size=1.2) + ylim(0,quantile(data_analysis2$Visits, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__+1), ymax=(upper__+1)), linetype=2, alpha=0.1,fill="darkblue") + ylab("Number of visits")+ ggtitle("A")

colnames(ce_2[[1]])[2] <- "Interaction"

Visits_PC2 <- ggplot(ce_2[[1]], aes(x = PC2, y = (estimate__+1))) + geom_point(data = data_analysis2,aes(x = PC2, y = Visits),
  size = 1.2, alpha=0.9) + geom_line(size=1.2) + ylim(0,quantile(data_analysis2$Visits, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__+1), ymax=(upper__+1)), linetype=2, alpha=0.1) + ylab("Number of visits") + ggtitle("B")

Specialization_PC1 <- ggplot(spe_1[[1]], aes(x = PC1, y = (estimate__))) + geom_point(data = data_analysis2,aes(x = PC1, y = d),
                      size = 1.2, alpha=0.9) + geom_line(size=1.2)  +theme_ms()+
                      geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.1) + ylab("Specilization (d')") + ggtitle("C")

Specialization_PC2 <- ggplot(spe_2[[1]], aes(x = PC2, y = (estimate__))) + geom_point(data = data_analysis2,aes(x = PC2, y = d),
                      size = 1.2, alpha=0.9) + geom_line(size=1.2)  +theme_ms()+
                      geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.1) + ylab("Specilization (d')") + ggtitle("D")

plot_grid(Visits_PC1,Visits_PC2,Specialization_PC1,Specialization_PC2)

```