### 2021-04-07 Supplementary_Material/Table_1_List_of_Studies.Rmd
Table s1 with all studies and information about the type of networks that contains

---
title: ""
output:
  pdf_document: default
  html_document: default
---
\begin{landscape}

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE,}
setwd("~/R_Projects/Reproductive Traits")
library(kableExtra)

#Create data.frame with references

`Number of networks` <- c("6", "2", "3", "1", "1", "1", "1", "2", "1", "1", "8", "16", "6", "2", "4", "1", "3", "1", "1", "1", "1", "1", "1", "1", "3", "1",
                          "1", "1")

`Network structure` <- c(rep("Web", 22), rep("Metaweb",6))

`Network type` <- c(rep("Weighted",  18), rep("Unweighted",  4), "Weighted", "Unweighted", "Unweighted", "Unweighted", "Unweighted", "Weighted")

`First author` <- c("Bartomeus", "Dicks", "Dupont", "Elberling", "Fang", "Inouye", "Lundgren", "Olesen", "Small", "Souza", "Kaiser-Bunbury", "Bartomeus", 
                    "Kaiser-Bunbury", "Kaiser-Bunbury", "Peralta", "Burkle", "Arroyo-Correa", "Bundgaard", "Bek", "Kato", "Kevan", "Ramirez",
                    "Inouye","McMullen", "Primack", "Ramirez", "Robertson", "Traveset")

Year <- c(2008,2002,2003,1999,2008,1988, 2005,2002,1976,2017,2017,2008,2011,2010,2006,2013,2019,2003,2006,2000,1970,1989,1990, 1993, 1983, 1992, 1929, 2013)

Country <- c("Spain", "England", "Denmark", "Sweden", "China", "Australia", "Greenland", "Mauritius and Azores", "Canada", "Brasil", "Seychelles", "Spain", "Seychelles",
"Mauritius", "Argentina", "USA", "New Zealand", "Denmark",  "Denmark", "Japan", "Canada", "Venezuela", "Japan", "Ecuador", "New Zealand", "Venezuela",
"USA", "Ecuador")

DOI <- c("https://doi.org/10.1007/s00442-007-0946-1", "https://doi.org/10.1046/j.0021-8790.2001.00572.x", "https://doi.org/10.1111/j.1365-2656.2008.01501.x",
"https://doi.org/10.1111/j.1600-0587.1999.tb00507.x", "https://doi.org/10.1111/1749-4877.12190", "https://doi.org/10.1111/j.1442-9993.1988.tb00968.x",
"https://doi.org/10.1657/1523-0430(2005)037[0514:TDAHCW]2.0.CO;2", "https://doi.org/10.1046/j.1472-4642.2002.00148.x",
"/13960/t4km08d21", "https://doi.org/10.1111/1365-2745.12978", "https://doi.org/10.1038/nature21071", "https://github.com/ibartomeus/BeeFunData",
"https://doi.org/10.1111/j.1365-2745.2010.01732.x", "https://doi.org/10.1016/j.ppees.2009.04.001", "https://doi.org/10.1111/ele.13510",
"https://doi.org/10.1126/science.1232728", "https://doi.org/10.1111/1365-2745.13332", "Unpublished, Master thesis", "Unpublished, Master thesis", "http://hdl.handle.net/2433/156116",
"https://doi.org/10.2307/2258569", "https://doi.org/10.2307/2388282", "http://hdl.handle.net/2433/156099", "https://biostor.org/reference/244737", "https://doi.org/10.1080/0028825X.1983.10428561",
"https://doi.org/10.1111/j.1095-8339.1992.tb00294.x", "https://doi.org/10.5962/bhl.title.11538", "https://doi.org/10.1098/rspb.2012.3040")

references <- data.frame(`First author`, `Number of networks`, `Network type`, `Network structure`, Year, Country, DOI)

colnames(references) <- c("First author", "Network N°", "Type", "Structure", "Year", "Country", "DOI")

kable(references, "latex", longtable = T, booktabs = T,linesep = "\\addlinespace") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 9, full_width=F)
```

\end{landscape}