### 2021-04-28 Main_Figures/Analysis_1_Separated.Rmd
plot 4 of the paper, not super fan of it but good for now

---
output:
  pdf_document: default
  html_document: default
classoption: landscape

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=28}

#LOAD LIBRARIES 
library(phytools)
library(ape) #for phylogenetic distance
library(dplyr) #data processing
library(rtrees) #for phylogenetic distancelibrary(MASS)
library(reshape2)
library(viridis) #COLOUR GGPLOT
library(MASS)
library(ggplot2)
library(broman) #crayon colours
library(tidyverse)
library(patchwork) #panel of plots
library(brms)
library(cmdstanr)

#Load theme for publication

theme_ms <- function(base_size=14, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(2.2)),
           axis.text=element_text(size = rel(2), color = "black"),
           legend.title=element_text(face="bold",size=rel(2)),
           legend.text=element_text(size=rel(1.4)),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(1.8, 'lines'),
           panel.border=element_rect(color="black",size=2.4),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank(),
           plot.title = element_text(hjust = 0.5,size=rel(2.75),face="bold"),
           plot.subtitle = element_text(face="bold",size=rel(2))
     ))
}

setwd("~/Dropbox/PhD/R/Chapter_2") #DROPBOX, files too large for github

#Plot nicely PC1
analysis_1_just_bees <- readRDS("results_analysis_1_just_bees.rds")
dat_analysis <- readRDS("dat_analysis_results_analysis_1_just_bees.rds")

ce_pc1 <- conditional_effects(analysis_1_just_bees, effects = "PC1",points=T) 
ce_pc1[[1]]$guild <- "bees"

p1 <- ggplot(ce_pc1[[1]], aes(x = -PC1, y = (estimate__+1))) + geom_point(data = dat_analysis,
  aes(x = -PC1, y = Interaction),size = 2.5, alpha=0.5, color="orange") + geom_line(size=1.4, color="orange") + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("Number of visits")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC1", subtitle = "(a)")

#Plot nicely PC2
ce_pc2 <- conditional_effects(analysis_1_just_bees, effects = "PC2",points=T) 
ce_pc1[[1]]$guild <- "bees"

p2 <- ggplot(ce_pc2[[1]], aes(x = -PC2, y = (estimate__+1))) + geom_point(data = dat_analysis,
  aes(x = -PC2, y = Interaction),size = 2.5, alpha=0.5, color="orange") + geom_line(size=1.4, color="orange") + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC2", subtitle = "(b)")
#Plot nicely PC3
ce_pc3 <- conditional_effects(analysis_1_just_bees, effects = "PC3",points=T)
ce_pc1[[1]]$guild <- "bees"

p3 <- ggplot(ce_pc3[[1]], aes(x = -PC3, y = (estimate__+1))) + geom_point(data = dat_analysis,
  aes(x = -PC3, y = Interaction),size = 2.5, alpha=0.5, color="orange") + geom_line(size=1.4, color="orange") + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC3", subtitle = "(c)")

#Plot nicely PC1
analysis_1_no_bees <- readRDS("results_analysis_1_no_bees.rds")
dat_analysis <- readRDS("dat_analysis_results_analysis_1_no_bees.rds")

ce_pc1 <- conditional_effects(analysis_1_no_bees, effects = "PC1:guild",points=T)

pp1 <- ggplot(ce_pc1[[1]], aes(x = -PC1, y = (estimate__+1), group=guild, colour=guild)) + geom_point(data = dat_analysis,
  aes(x = -PC1, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("Number of visits")+ xlab("High Flower N.    Large flowers")+
  theme_ms() + theme(legend.position = "none",plot.title = element_text(size=rel(2),hjust = 0,face="bold")) + scale_color_manual(name="Functional groups",labels=c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera",
   "Non-syrphids-diptera", "Syrphids"),values=c("orange", "black", "limegreen","#E7298A", "cyan4","blueviolet"),
    limits = c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera","Non-syrphids-diptera", "Syrphids")) + ggtitle("(d)")

ce_pc2 <- conditional_effects(analysis_1_no_bees, effects = "PC2:guild",points=T)

pp2 <- ggplot(ce_pc2[[1]], aes(x = -PC2, y = (estimate__+1), group=guild, colour=guild)) + geom_point(data = dat_analysis,
  aes(x = -PC2, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("Low poll. depend. High poll. depend.")+
  theme_ms() + theme(legend.position = "none",plot.title = element_text(size=rel(2),hjust = 0,face="bold")) + scale_color_manual(name="Functional groups",labels=c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera",
   "Non-syrphids-diptera", "Syrphids"),values=c("orange", "black", "limegreen","#E7298A", "cyan4","blueviolet"),
    limits = c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera","Non-syrphids-diptera", "Syrphids")) + ggtitle("(e)")

#Plot nicely PC3
ce_pc3 <- conditional_effects(analysis_1_no_bees, effects = "PC3:guild",points=T)

pp3 <- ggplot(ce_pc3[[1]], aes(x = -PC3, y = (estimate__+1), group=guild, colour=guild)) + geom_point(data = dat_analysis,
  aes(x = -PC3, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("Short styles    Long styles")+
  theme_ms() +theme(legend.position = "none",plot.title = element_text(size=rel(2),hjust = 0,face="bold")) + scale_color_manual(name="Functional groups",labels=c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera",
    "Non-syrphids-diptera", "Syrphids"),values=c("orange", "black", "limegreen","#E7298A", "cyan4","blueviolet"),
    limits = c("Bees","Coleoptera", "Lepidoptera", "Non-bee-Hymenoptera",
               "Non-syrphids-diptera", "Syrphids")) + ggtitle("(f)")

library(patchwork)

combined <- p1 + p2 + p3 + pp1 + pp2 + pp3 & theme(legend.position = "right")
combined + plot_layout(guides = "collect",ncol = 3)

```