### 2021-04-29 Supplementary_Material/Missing_data.Rmd

---
output:
  pdf_document: default
  html_document: default

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE,dpi=2000}

#Set working directory
setwd("~/R_Projects/Reproductive Traits")

########################################################################################################################################################
#SCRIPT TO CHECK PATTERNS OF MISSING VALUES
########################################################################################################################################################
#1) READ TRAIT DATA

#2) CLEAN DATA

#3) EXPLORE PATTERNS OF MISSING DATA
########################################################################################################################################################
#LOAD LIBRARIES
library(readxl) #read excel file (trait data)
library(dplyr) #data manipulation
library(missMDA) #For missing values
library(FactoMineR) #Produce pca biplot with individuals and variables
library(factoextra)
library(ggpubr) 
library(missForest) #random forst imputation
library(rtrees) #for phylogenetic distancelibrary(MASS)
library(ape)
library(PVR) #calculate eigen vectors and add them as a column in the imputation, seems to improve output
library(visdat)
library(tidyverse)

########################################################################################################################################################
#1) READ TRAIT DATA
########################################################################################################################################################
#load data
trait_data <- read_excel("Data/Trait_data_raw/Trait_data_final.xlsx",na = "NA")
########################################################################################################################################################
#2) A) CLEAN DATA AND B) RENAME LEVELS FOR IMPUTATION 
########################################################################################################################################################
#####
#A) CLEAN DATA
#####
#select just filled rows
trait_data_1 <- trait_data[1:1712,] #It may be a more elegant way but this does the job, 1712 rows of data

#str(trait_data_1)
#filter data, select species with flower level info and capitulum
trait_filtered <- filter(trait_data_1, Info_level == "flower" |  Info_level == "capitulum")
#levels(as.factor(trait_filtered$Info_level)) #checking levels
#remove NA's and duplicated species|Some spp are repeated once the spp names are standardize with taxize
trait_filtered$Species_all[trait_filtered$Species_all=="NA"]<-NA 
trait_filtered_1 <- trait_filtered[!is.na(trait_filtered$Species_all),]
trait_filtered_2 <- trait_filtered_1[!duplicated(trait_filtered_1$Species_all),]
#str(trait_filtered_2)

#Select columns to work with
t <- trait_filtered_2[c("Species_geonet","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul")]

####################################################################################################
#2) CLEAN DATA. I REPEAT THE SAME PROCEDURE OF THE DATA IMPUTATION TO GET THE SAME NUMBER OF SPECIES
####################################################################################################
#Before data imputation I need to calculate the phylo of the species
#remove not found species that do not run with get tree
cols.num <- c("Family_all","Genus_all","Species_all")
t[cols.num] <- sapply(t[cols.num],as.character)
t$Species_all <- gsub("Species_all_", "", t$Species_all)

t <- t[!t$Species_all == "Pinus luchuensis", ]   # remove gymnosperm species

#remove species that are not until species level
#ALL THE SPECIES WITH SP. ARE DELETD
t$Species_geonet <- gsub("M_PL_.*","",t$Species_geonet) #subsitute partial string for nothing
t$Species_geonet <- gsub(" $","", t$Species_geonet, perl=T) #remove trailing spaces
t$Species_geonet <-  sub('sp.1$', 'sp.', t$Species_geonet)#PREPARE ALL SP TO SP.
t$Species_geonet <- sub('sp.2$', 'sp.', t$Species_geonet)#PREPARE ALL SP TO SP.
t$Species_geonet <- sub('sp$', 'sp.', t$Species_geonet) #PREPARE ALL SP TO SP.
t$Species_geonet <- sub("sp.$", "DELETE", t$Species_geonet) #change all sp. for DELETE
t <- t[- grep("DELETE",  t$Species_geonet),] #remove species with "DELETE"
#CHECK LEVELS
#levels(factor(t$Species_all))

#Make these NA's as NA
t$Species_all[t$Species_all=="NA"] <- NA
t$Family_all[t$Family_all=="NA"] <- NA
t$Genus_all[t$Genus_all=="NA"] <- NA
#remove NA's
t <- t[!is.na(t$Family_all),]
t <- t[!is.na(t$Species_all),]
t <- t[!is.na(t$Genus_all),]

#nrow(t)

#Select just columns with traits
t <- t[c("Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul")]

#################################################
#3)SUMMARY OF THE DATA|ALL COLS CONSIDERED!
#################################################
##Number of filled cells
#nrow(t) * 16 - sum(is.na(t))#[1] 20737
##Dimension of the dataset
#nrow(t) * 16#[1] 24096
##Number of missing values
#sum(is.na(t))#[1] 3359
##Percentage of filled cells
#sum(!is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100#[1] 86.87993
##Percentage of missing values
#sum(is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100#[1] 13.12007
#
#

#Function to specify number of decimals
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))

no_missing <- specify_decimal(sum(!is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100,2)
missing <- specify_decimal(sum(is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100,2)

colnames(t) <- c("Breeding system","Compatibility system",
         "Autonomous selfing qualitative","Autonomous selfing quantitative", "Flower shape", "Flower symmetry", 
         "Flower number per plant", "Floral display width", "Flower width", "Flower length","Style length", "Ovule number", 
         "Life form", "Lifespan","Plant height","Nectar qualitative","Nectar quantitative")

missing.values <- t %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <-
  (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
  ggplot() +
  geom_bar(aes(x = reorder(key, desc(pct)), 
               y = pct, fill=isna), 
           stat = 'identity', alpha=0.8) +
  scale_x_discrete(limits = levels) + theme_bw()+
  scale_fill_manual(name = "", 
                    values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
  coord_flip() +
  labs(title = "Percentage of missing values", x =
         'Variables', y = "% of missing values")

percentage.plot

```