### 2021-05-11 Main_Figures/Traits_Table.Rmd
.

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=12, fig.width=18}

library(kableExtra)
library(dplyr)

Trait <- c("Plant height (m)", " ", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovule number/flower",
           "Flowers per plant", "Nectar $(\\mu l)$", "Nectar (mg)", "Nectar concentration (%)", "Pollen grains per flower", "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Nectar", "Autonomous selfing", "Compatibility system", "Breeding system", " ", " ", " ", " ", " ")

T.cat <- c("Short-lived, perennial", "Herb, shrub, tree", "Brush, campanulate, capitulum, open, papilionaceous, tube", "Actinomorphic, zygomorphic",
           "Presence/absence", "None, low, medium, high", "Incompatible, partially compatible, compatible", "Hermaphrodite, monoecious, dioecious"," ", " ", " ", " ", " ")

Type <- c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Floral","Floral","Floral","Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Floral", "Reproductive", "Reproductive","Reproductive","","","","")

dat <- data.frame(Type, Trait,Type_1, Trait1, T.cat)

colnames(dat) <- c("Trait Type","Traits","Trait Type", "Traits", "Categories")

#dat%>%
#kbl("latex", longtable = T, booktabs = T,linesep = "\\addlinespace")

#%>%

# add_header_above(c("Quantitative traits" = 2,"Qualitative traits" = 3))

#%>%  column_spec(c(1,3), bold = T,width="2em",extra_css="transform: rotate(-90deg);") %>%
 # collapse_rows(columns = c(1,3), valign = "middle")

kable(dat,"latex", longtable = T, booktabs = T,linesep = "\\addlinespace") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "left") %>% collapse_rows(columns = c(1,3), valign = "middle")

```