### 2021-05-11 Main_Figures/Regression_Tree_Nectar_Pollen.Rmd
prepare table at home

---
title: ""
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18}

########################################################################################################################################################
#SCRIPT TO CALCULATE THE REGRESSION TREE OF NECTAR AND PLOT IT
########################################################################################################################################################
#READ NETWORK VISITATION DATA (VISITS AGGREGATED PER PLANT SPECIES)

library(rpart.plot)
library(rpart)

setwd("~/R_Projects/Reproductive Traits")

long_d <- read.csv("Data/Csv/long_format_quantitative_networks.csv", row.names = 1) #quantitative network data|weighted by frequency of visits per plant species

#Select data with interaction greater than 0
long_d_1 <- long_d[long_d$Interaction>0,]

#Remove NA'S
long_d_2 <- long_d_1[!is.na(long_d_1$guild),] #I do it by guild because just these 6 guilds are named in this column

#check levels
levels(factor(long_d_2$guild))
########################################################################################################################################################
#NOW READ NECTAR DATA
forest_data <- read.csv("Data/Csv/nectar_pollen_subset_imputed_trait_data.csv")
colnames(forest_data)[1] <- "Plant_species"
########################################################################################################################################################
#MERGE DATA
data <- merge(long_d_2, forest_data, by="Plant_species")

#SELECT COLUMNS WITH QUANTITATIVE INFORMATION
data_1 <- data[c("Interaction","Autonomous_selfing_level_fruit_set","Corolla_diameter_mean","Corolla_length_mean",
                 "Flowers_per_plant","Style_length","Ovule_number","Plant_height_mean_m","Nectar_concentration",
                 "Nectar_mg","Nectar_ul","Pollen_per_flower")]

#Set colnames for plotting
colnames(data_1) <- c("Interactions", "Aut. selfing", "Flower width","Flower length" , "Flowers per plant","Style length (mm)",
                      "Ovule number","Plant height (m)", "Nectar concentration (%)","Mg of nectar","Microlitres of nectar","Pollen grains per flower")
########################################################################################################################################################

# Create a decision tree model
tree <- rpart(Interactions~., data=data_1, cp=.007)
# Visualize the decision tree with rpart.plot
rpart.plot(tree, box.palette="GnOr",tweak = 1.20)

```