### 2021-05-11 Main_Figures/Traits_Table.Rmd

---
title: ""
output:
  pdf_document: default
  html_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=12, fig.width=18}

library(kableExtra)

library(dplyr)

Trait <- c("Plant height (m)", " ", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovule number/flower",
           "Flowers/plant", "Nectar ($\\mu$l)", "Nectar (mg)", "Nectar concentration ($\\%$)", "Pollen grains/flower", "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Nectar", "Autonomous selfing", "Compatibility system", "Breeding system", " ", " ", " ", " ", " ")

T.cat <- c("Short-lived, perennial", "Herb, shrub, tree", "Brush, campanulate, capitulum, open, papilionaceous, tube", "Actinomorphic, zygomorphic",
           "Presence/absence", "None, low, medium, high", "Self-incompatible, partially self-compatible, self-compatible", "Hermaphrodite, monoecious, dioecious"," ", " ", " ", " ", " ")

Type <- c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Floral","Floral","Floral","Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Floral", "Reproductive", "Reproductive", "Reproductive","","","","","")

dat <- data.frame(Type, Trait,Type_1, Trait1, T.cat)

colnames(dat) <- c("Trait type","Traits","Trait type", "Traits", "Categories")

#dat%>%
#kbl("latex", longtable = T, booktabs = T,linesep = "\\addlinespace")

#%>%

# add_header_above(c("Quantitative traits" = 2,"Qualitative traits" = 3))

#%>%  column_spec(c(1,3), bold = T,width="2em",extra_css="transform: rotate(-90deg);") %>%
 # collapse_rows(columns = c(1,3), valign = "middle")

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE) %>%
kable_styling(full_width = T) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 3),bold=T)  %>%
    column_spec(c(1,3), bold = T) %>% row_spec(0, bold=T)

```