### 2021-05-11 Supplementary_Material/Table_1_List_of_Studies.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\begin{landscape}

```
## [1] "latex"
```

\begingroup\fontsize{12}{14}\selectfont

\begin{longtable}[l]{lrlll}
\toprule
First author & Year & Web N. & Network type & DOI\\
\midrule
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\toprule
First author & Year & Web N. & Network type & DOI\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{Arroyo-Correa} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{3} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1111/1365-2745.13332}\\
\addlinespace
Bartomeus & 2008 & 6 & Weighted web & https://doi.org/10.1007/s00442-007-0946-1\\
\addlinespace
\cellcolor{gray!6}{Bartomeus} & \cellcolor{gray!6}{2008} & \cellcolor{gray!6}{16} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://github.com/ibartomeus/BeeFunData}\\
\addlinespace
Bek & 2006 & 1 & Unweighted web & Unpublished, Master thesis\\
\addlinespace
\cellcolor{gray!6}{Bundgaard} & \cellcolor{gray!6}{2003} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{Unpublished, Master thesis}\\
\addlinespace
Burkle & 2013 & 1 & Weighted web & https://doi.org/10.1126/science.1232728\\
\addlinespace
\cellcolor{gray!6}{Dicks} & \cellcolor{gray!6}{2002} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1046/j.0021-8790.2001.00572.x}\\
\addlinespace
Dupont & 2003 & 3 & Weighted web & https://doi.org/10.1111/j.1365-2656.2008.01501.x\\
\addlinespace
\cellcolor{gray!6}{Elberling} & \cellcolor{gray!6}{1999} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1111/j.1600-0587.1999.tb00507.x}\\
\addlinespace
Fang & 2008 & 1 & Weighted web & https://doi.org/10.1111/1749-4877.12190\\
\addlinespace
\cellcolor{gray!6}{Inouye} & \cellcolor{gray!6}{1988} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1111/j.1442-9993.1988.tb00968.x}\\
\addlinespace
Inouye & 1990 & 1 & Weighted metaweb & http://hdl.handle.net/2433/156099\\
\addlinespace
\cellcolor{gray!6}{Kaiser-Bunbury} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{8} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1038/nature21071}\\
\addlinespace
Kaiser-Bunbury & 2011 & 6 & Weighted web & https://doi.org/10.1111/j.1365-2745.2010.01732.x\\
\addlinespace
\cellcolor{gray!6}{Kaiser-Bunbury} & \cellcolor{gray!6}{2010} & \cellcolor{gray!6}{2} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1016/j.ppees.2009.04.001}\\
\addlinespace
Kato & 2000 & 1 & Unweighted web & http://hdl.handle.net/2433/156116\\
\addlinespace
\cellcolor{gray!6}{Kevan} & \cellcolor{gray!6}{1970} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Unweighted web} & \cellcolor{gray!6}{https://doi.org/10.2307/2258569}\\
\addlinespace
Lundgren & 2005 & 1 & Weighted web & https://doi.org/10.1657/1523-0430(2005)037[0514:TDAHCW]2.0.CO;2\\
\addlinespace
\cellcolor{gray!6}{McMullen} & \cellcolor{gray!6}{1993} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Unweighted metaweb} & \cellcolor{gray!6}{https://biostor.org/reference/244737}\\
\addlinespace
Olesen & 2002 & 2 & Weighted web & https://doi.org/10.1046/j.1472-4642.2002.00148.x\\
\addlinespace
\cellcolor{gray!6}{Peralta} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{4} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1111/ele.13510}\\
\addlinespace
Primack & 1983 & 3 & Unweighted metaweb & https://doi.org/10.1080/0028825X.1983.10428561\\
\addlinespace
\cellcolor{gray!6}{Ramirez} & \cellcolor{gray!6}{1989} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Unweighted web} & \cellcolor{gray!6}{https://doi.org/10.2307/2388282}\\
\addlinespace
Ramirez & 1992 & 1 & Unweighted metaweb & https://doi.org/10.1111/j.1095-8339.1992.tb00294.x\\
\addlinespace
\cellcolor{gray!6}{Robertson} & \cellcolor{gray!6}{1929} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Unweighted metaweb} & \cellcolor{gray!6}{https://doi.org/10.5962/bhl.title.11538}\\
\addlinespace
Small & 1976 & 1 & Weighted web & /13960/t4km08d21\\
\addlinespace
\cellcolor{gray!6}{Souza} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{Weighted web} & \cellcolor{gray!6}{https://doi.org/10.1111/1365-2745.12978}\\
\addlinespace
Traveset & 2013 & 1 & Weighted metaweb & https://doi.org/10.1098/rspb.2012.3040\\*
\end{longtable}
\endgroup{}

\end{landscape}

\end{document}