### 2021-05-17 Supplementary_Material/Analysis_2_Hymenoptera_Families_Separated.Rmd

---
output:
  pdf_document: default
  html_document: default
classoption: landscape

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=28}

#LOAD LIBRARIES 
library(phytools)
library(ape) #for phylogenetic distance
library(dplyr) #data processing
library(rtrees) #for phylogenetic distancelibrary(MASS)
library(reshape2)
library(viridis) #COLOUR GGPLOT
library(MASS)
library(ggplot2)
library(broman) #crayon colours
library(tidyverse)
library(patchwork) #panel of plots
library(brms)
library(cmdstanr)

#Load theme for publication

theme_ms <- function(base_size=14, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(2.2)),
           axis.text=element_text(size = rel(2), color = "black"),
           legend.title=element_text(face="bold",size=rel(2)),
           legend.text=element_text(size=rel(1.4)),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(1.8, 'lines'),
           panel.border=element_rect(color="black",size=2.4),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank()
     ))
}

setwd("~/Dropbox/PhD/R/Chapter_2") #DROPBOX, files too large for github

#Plot nicely PC1
analysis_1 <- readRDS("results_analysis_1_bee_families.rds")
dat_analysis <- readRDS("dat_analysis_results_analysis_1_bee_families.rds")

ce_pc1 <- conditional_effects(analysis_1, effects = "PC1:bee_family",points=T)

a1 <- ggplot(ce_pc1[[1]], aes(x = -PC1, y = (estimate__+1), group=bee_family, colour=bee_family)) + geom_point(data = dat_analysis,
  aes(x = -PC1, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("Number of visits")+ xlab("PC1")+
  theme_ms() + theme(legend.position = "none") +  scale_color_manual(name="Functional groups",labels=c("Andrenidae","Apidae", "Colletidae", "Halictidae",
   "Megachilidae", "All families"),values=c("blueviolet", "black", "limegreen","#E7298A", "cyan4","orange"),
    limits = c("Andrenidae","Apidae", "Colletidae", "Halictidae","Megachilidae", "All families"))+ggtitle("(a)") + theme(plot.title = element_text(size = 28, face = "bold"))
#Plot nicely PC2
ce_pc2 <- conditional_effects(analysis_1, effects = "PC2:bee_family",points=T)

a2 <- ggplot(ce_pc2[[1]], aes(x = -PC2, y = (estimate__+1), group=bee_family, colour=bee_family)) + geom_point(data = dat_analysis,
  aes(x = -PC2, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("PC2")+
  theme_ms() + theme(legend.position = "none") +  scale_color_manual(name="Functional groups",labels=c("Andrenidae","Apidae", "Colletidae", "Halictidae",
   "Megachilidae", "All families"),values=c("blueviolet", "black", "limegreen","#E7298A", "cyan4","orange"),
    limits = c("Andrenidae","Apidae", "Colletidae", "Halictidae","Megachilidae", "All families"))+ggtitle("(b)") + theme(plot.title = element_text(size = 28, face = "bold"))

#Plot nicely PC3
ce_pc3 <- conditional_effects(analysis_1, effects = "PC3:bee_family",points=T)

a3 <- ggplot(ce_pc3[[1]], aes(x = -PC3, y = (estimate__+1), group=bee_family, colour=bee_family)) + geom_point(data = dat_analysis,
  aes(x = -PC3, y = Interaction),size = 2.5, alpha=0.5) + geom_line(size=1.4) + 
  ylim(0,quantile(dat_analysis$Interaction, 0.95)) + ylab("")+ xlab("PC3")+
  theme_ms() + theme(legend.position = "none") +  scale_color_manual(name="Functional groups",labels=c("Andrenidae","Apidae", "Colletidae", "Halictidae",
   "Megachilidae", "All families"),values=c("blueviolet", "black", "limegreen","#E7298A", "cyan4","orange"),
    limits = c("Andrenidae","Apidae", "Colletidae", "Halictidae","Megachilidae", "All families"))+ggtitle("(c)") + theme(plot.title = element_text(size = 28, face = "bold"))

library(patchwork)

combined <- a1 + a2 + a3 & theme(legend.position = "right")
combined + plot_layout(guides = "collect", ncol=3)

```