### 2021-05-20 Supplementary_Material/Species_Level_Metric_Pc's.Rmd
Metrics~Pc's added to supplementary

---
output:
  pdf_document: default
  html_document: default
classoption: landscape

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=20, fig.width=28}

library(brms) #modelling
library(cmdstanr) #modelling
library(cowplot) #panel of plots
library(ggplot2) #plotting
library(patchwork)

# Theme for publication
theme_ms <- function(base_size=12, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(2)),
           axis.text=element_text(size = rel(2.5), color = "black"),
           legend.title=element_text(face="bold"),
           legend.text=element_text(),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(0.4, 'lines'),
           panel.border=element_rect(color="black",size=2),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank()
     ))
}
#########

setwd("~/Dropbox/PhD/R/Chapter_2") #DROPBOX, files too large for github

#read data for plotting
data_analysis_2 <- readRDS("data_analysis_2.rds")

#INTERACTION FREQUENCY
Visits_PCs <- readRDS("results_analysis_2_Visits.rds") #load model

ce_1 <- conditional_effects(Visits_PCs, effects = "PC1",points=T) 
colnames(ce_1[[1]])[3] <- "Interaction"

Visits_PC1 <- ggplot(ce_1[[1]], aes(x = -PC1, y = (estimate__+1))) + geom_point(data = data_analysis_2,aes(x = -PC1, y = Visits),
  size = 2.5, alpha=0.65, colour="goldenrod3") + geom_line(colour="black",size=1.5) + ylim(0,quantile(data_analysis_2$Visits, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__+1), ymax=(upper__+1)), linetype=2, alpha=0.1,fill="black") + ylab("Sum of visits") + xlab("PC1")

ce_2 <- conditional_effects(Visits_PCs, effects = "PC2",points=T) 
colnames(ce_2[[1]])[3] <- "Interaction"

Visits_PC2 <- ggplot(ce_2[[1]], aes(x = -PC2, y = (estimate__+1))) + geom_point(data = data_analysis_2,aes(x = -PC2, y = Visits),
  size = 2.5, alpha=0.65, colour="goldenrod3") + geom_line(colour="black",size=1.5) + ylim(0,quantile(data_analysis_2$Visits, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__+1), ymax=(upper__+1)), linetype=2, alpha=0.1,fill="black") + ylab("Sum of visits") + xlab("PC2")

ce_3 <- conditional_effects(Visits_PCs, effects = "PC3",points=T) 
colnames(ce_3[[1]])[3] <- "Interaction"

Visits_PC3 <- ggplot(ce_3[[1]], aes(x = -PC3, y = (estimate__+1))) + geom_point(data = data_analysis_2,aes(x = -PC3, y = Visits),
  size = 2.5, alpha=0.65, colour="goldenrod3") + geom_line(colour="black",size=1.5) + ylim(0,quantile(data_analysis_2$Visits, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__+1), ymax=(upper__+1)), linetype=2, alpha=0.1,fill="black") + ylab("Sum of visits") + xlab("PC3")

#NORMALIZED DEGREE
normalised.degree_PCs <- readRDS("results_analysis_2_normalised.degree_PCs.rds") #load model

ce_1 <- conditional_effects(normalised.degree_PCs, effects = "PC1",points=T)

normalised.degree_PC1 <- ggplot(ce_1[[1]], aes(x = -PC1, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC1, y = normalised.degree),
  size = 2.5, alpha=0.65, colour="aquamarine4") + geom_line(colour="black",size=1.5)  +theme_ms()+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.25,fill="black") + ylab("Normalized degree") + xlab("PC1")

ce_2 <- conditional_effects(normalised.degree_PCs, effects = "PC2",points=T)

normalised.degree_PC2 <- ggplot(ce_2[[1]], aes(x = -PC2, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC2, y = normalised.degree),
  size = 2.5, alpha=0.65,colour="aquamarine4") + geom_line(colour="black",size=1.5) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.25,fill="black") + ylab("Normalized degree") + xlab("PC2")

ce_3 <- conditional_effects(normalised.degree_PCs, effects = "PC3",points=T)

normalised.degree_PC3 <- ggplot(ce_3[[1]], aes(x = -PC3, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC3, y = normalised.degree),
  size = 2.5, alpha=0.65,colour="aquamarine4") + geom_line(colour="black",size=1.5)  +theme_ms()+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.25,fill="black") + ylab("Normalized degree") + xlab("PC3")

#SPECIALIZATION
Specialization_PCs <- readRDS("results_analysis_2_Specialization_PCs.rds") #load model

ce_1 <- conditional_effects(Specialization_PCs, effects = "PC1",points=T) 
colnames(ce_1[[1]])[3] <- "Interaction"

Specialization_PC1 <- ggplot(ce_1[[1]], aes(x = -PC1, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC1, y = d),
  size = 2.5, alpha=0.65, colour="darkorange3") + geom_line(colour="black",size=1.5) + ylim(0,quantile(data_analysis_2$d, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.1,fill="black") + ylab("Specialization (d')") + xlab("PC1")

ce_2 <- conditional_effects(Specialization_PCs, effects = "PC2",points=T) 
colnames(ce_2[[1]])[3] <- "Interaction"

Specialization_PC2 <- ggplot(ce_2[[1]], aes(x = -PC2, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC2, y = d),
  size = 2.5, alpha=0.65, colour="darkorange3") + geom_line(colour="black",size=1.5) + ylim(0,quantile(data_analysis_2$d, 0.95)) +theme_ms()+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.1,fill="black") + ylab("Specialization (d')") + xlab("PC2")

ce_3 <- conditional_effects(Specialization_PCs, effects = "PC3",points=T) 
colnames(ce_3[[1]])[3] <- "Interaction"

Specialization_PC3 <- ggplot(ce_3[[1]], aes(x = -PC3, y = (estimate__))) + geom_point(data = data_analysis_2,aes(x = -PC3, y = d),
 size = 2.5, alpha=0.65, colour="darkorange3") + ylim(0,quantile(data_analysis_2$d, 0.95)) +theme_ms()+geom_line(colour="black",size=1.5)+
  geom_ribbon(aes(ymin=(lower__), ymax=(upper__)), linetype=2, alpha=0.1,fill="black") + ylab("Specialization (d')") + xlab("PC3")

Visits_PC1 + Visits_PC2 + Visits_PC3 + normalised.degree_PC1 + normalised.degree_PC2 + normalised.degree_PC3 +
  Specialization_PC1 + Specialization_PC2+ Specialization_PC3  + plot_layout(ncol = 3)

```