### 2021-09-19 Supplementary_Material/Apis_Non_Apis_Bees_Presence_Absence.Rmd
Update in figure and figure caption of supp mat

---
title: ""
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= ""}

library(reshape2)
library(viridis) #color GGPLOT
#library(MASS)
library(ggplot2)
library(broman) #crayon colors
#library(tidyverse)
library(patchwork) #panel of plots
library(brms)
#library(cmdstanr)

#Load theme for publication

theme_ms <- function(base_size=14, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
     theme(text=element_text(color="black"),
           axis.title=element_text( size = rel(2)),
           axis.text=element_text(size = rel(1.8), color = "black"),
           legend.title=element_text(face="bold",size=rel(2)),
           legend.text=element_text(size=rel(1.4)),
           legend.background=element_rect(fill="transparent"),
           legend.key.size = unit(1.8, 'lines'),
           panel.border=element_rect(color="black",size=2.4),
           panel.grid.minor.x =element_blank(),
           panel.grid.minor.y= element_blank(),
           panel.grid.major= element_blank(),
           plot.title = element_text(hjust = 0.5,size=rel(2.75),face="bold"),
           plot.subtitle = element_text(face="bold",size=rel(2))
     ))
}

#decimal places for ggplot

scaleFUN <- function(x) sprintf("%.1f", x)

#Prepare plot with presence absence and interaction frequency
setwd("~/Dropbox/PhD/R/Chapter_2") #DROPBOX, files too large for github
#Plot nicely PC1
analysis_1 <- readRDS("results_analysis_1_bee_apis_presence_absence.rds")
dat_analysis <- readRDS("dat_analysis_results_analysis_1_bee_apis_presence_absence.rds")

ce_pc1 <- conditional_effects(analysis_1, effects = "PC1",points=T)

pp1 <- ggplot(ce_pc1[[1]], aes(x = -PC1, y = (estimate__))) + geom_point(data = dat_analysis,
aes(x = -PC1, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5,  colour="orange")  + ylab("Interaction")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC1", subtitle = "(a)")+scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

ce_pc2 <- conditional_effects(analysis_1, effects = "PC2",points=T)

pp2 <- ggplot(ce_pc2[[1]], aes(x = -PC2, y = (estimate__))) + geom_point(data = dat_analysis,
aes(x = -PC2, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5, colour="orange")  + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC2", subtitle = "(b)") + scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

ce_pc3 <- conditional_effects(analysis_1, effects = "PC3",points=T)

pp3 <- ggplot(ce_pc3[[1]], aes(x = -PC3, y = (estimate__))) + geom_point(data = dat_analysis,
 aes(x = -PC3, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5,  colour="orange")  + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC3", subtitle = "(c)")+ scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

#Plot nicely PC1
analysis_1 <- readRDS("results_analysis_1_bee_non_apis_presence_absence.rds")
dat_analysis <- readRDS("dat_analysis_results_analysis_1_bee_non_apis_presence_absence.rds")

ce_pc1 <- conditional_effects(analysis_1, effects = "PC1",points=T)

p1 <- ggplot(ce_pc1[[1]], aes(x = -PC1, y = (estimate__))) + geom_point(data = dat_analysis,
aes(x = -PC1, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5,  colour="orange")  + ylab("Interaction")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC1", subtitle = "(a)")+scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

ce_pc2 <- conditional_effects(analysis_1, effects = "PC2",points=T)

p2 <- ggplot(ce_pc2[[1]], aes(x = -PC2, y = (estimate__))) + geom_point(data = dat_analysis,
aes(x = -PC2, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5,  colour="orange")  + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC2", subtitle = "(b)") + scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

ce_pc3 <- conditional_effects(analysis_1, effects = "PC3",points=T)

p3 <- ggplot(ce_pc3[[1]], aes(x = -PC3, y = (estimate__))) + geom_point(data = dat_analysis,
 aes(x = -PC3, y = Interaction),size = 2.5, alpha=0.4,  colour="orange") + geom_line(size=1.5,  colour="orange")  + ylab("")+ xlab("")+
  theme_ms() + theme(legend.position = "none") + ggtitle("PC3", subtitle = "(c)")+ scale_y_continuous(breaks = c(0,  1), labels=c("No", "Yes"))

combined <- pp1 + pp2 + pp3 + p1 + p2 + p3 & theme(legend.position = "bottom")
 combined + plot_layout(guides = "collect",ncol = 3)

```