### 2021-10-19 Manuscript/Manuscript.tex

\documentclass[12pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={ Trade-offs among plant reproductive traits determine interactions with floral visitors},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{lineno} % add 
\linenumbers % turns line numbering on

% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{titlesec}
\titlespacing{\title}{0pt}{\parskip}{-\parskip}
\usepackage[style=nature]{biblatex}
\usepackage[labelformat = empty]{caption}
\usepackage{pdflscape}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant
reproductive traits determine interactions with floral visitors}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{-2.2cm} \singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

\normalsize

\textbf{Plant life strategies are constrained by cost-benefit trade-offs
that determine plant form and function. However, despite recent advances
in the understanding of trade-offs for vegetative and physiological
traits, little is known about plant reproductive economics and how they
constrain plant life strategies and shape interactions with floral
visitors. Here, we investigate plant reproductive trade-offs and how
these reproductive compromises drive interactions with floral visitors
using a dataset of 17 reproductive traits for 1,506 plant species from
28 plant-pollinator studies across 18 countries. We tested whether a
plant's reproductive strategy predicts its interactions with floral
visitors and if the different reproductive traits predict the plant's
role within the pollination network. We found that over half of all
plant reproductive trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size, while the second axis indicated a pollinator dependency trade-off.
Plant reproductive trade-offs somewhat explained the presence or absence
of interactions with floral visitors, but not differences in visitation
rate. However, we did find important differences in the interaction
level among floral visitor guilds on the different axes of trait
variation. Finally, we found that plant size and floral rewards were the
most important traits in the understanding of the species network role.
Our results highlight the importance of plant reproductive trade-offs in
determining plant life strategies and plant-pollinator interactions in a
global context.}

\small
\vspace{-0.5cm}

\noindent\rule{\textwidth}{1pt}

\textsuperscript{1} School of Environmental and Rural Science,
University of New England, Armidale, New South Wales 2350, Australia.
\textsuperscript{2} Estación Biológica de Doñana (EBD-CSIC), E-41092
Seville, Spain. \textsuperscript{3} Department of Conservation
\textbar{} Te Papa Atawhai, Auckland, New Zealand. \textsuperscript{4}
Centre for Environmental and Climate Science, Lund University,
Sölvegatan 37, S-223 62 Lund, Sweden. \texttt{*} e-mail:
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}

\doublespacing
\vspace{5mm} \normalsize

Despite the astonishing diversity of floral structures among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}}
and their importance in shaping plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}},
a unified framework that explores plant reproductive compromises is
currently lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{5}}.
In addition, macroecological studies that investigate plant reproductive
traits are
scarce\textsuperscript{\protect\hyperlink{ref-baude2016}{6}--\protect\hyperlink{ref-moeller2017}{9}}
and consequently, there is poor understanding of how reproductive traits
drive interactions with floral visitors at large
scales\textsuperscript{\protect\hyperlink{ref-sargent2008}{10}--\protect\hyperlink{ref-ruger2018}{13}}.
Linking the position in the trait-space with the different pollinator
groups could help to improve our understanding of plant-pollinator
associations\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}}.
Further, there is increasing interest in understanding drivers of
plant-pollinator interactions using trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{15}}
and trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-bartomeus2016}{17}}.
However, reproductive traits have been overlooked beyond highly
specialised pollination
systems\textsuperscript{\protect\hyperlink{ref-dellinger2020}{4}}
despite the apparent generalist nature of plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-waser1996}{18},\protect\hyperlink{ref-olesen2002}{19}}.
In addition, it is unclear how specific plant reproductive biology
traits influence plant-pollinator
interactions\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-devaux2014}{21}}.

Species can optimise their fitness through various life-history traits,
yet trade-offs among those traits constrain the range of potential
strategies that a species can use. With the recent availability of large
trait databases (e.g.,
TRY\textsuperscript{\protect\hyperlink{ref-kattge2011}{22}} and
COMPADRE\textsuperscript{\protect\hyperlink{ref-salguero2015}{23}}),
plant ecological strategies are being increasingly examined, and are
facilitating the identification of global patterns and constraints in
plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{12},\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}}.
However, most studies have focused on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{26}},
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{27}}, or
root\textsuperscript{\protect\hyperlink{ref-laughlin2021}{28}}
trade-offs with little or no attention given to reproductive
traits\textsuperscript{\protect\hyperlink{ref-roddy2021}{5},\protect\hyperlink{ref-evojtko2020}{29}}
which are critical to plant life strategies that shape interactions with
pollinators and ultimately determine plant reproductive success. For
instance, short lived versus perennial species tend to have low versus
high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{9},\protect\hyperlink{ref-barrett2003}{30}}
and outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and its role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{32},\protect\hyperlink{ref-rodger2021}{33}}.
However, it is still unknown to what extent these different reproductive
compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and
plant-pollinator network
properties\textsuperscript{\protect\hyperlink{ref-bartomeus2013}{34}--\protect\hyperlink{ref-rowe2020}{36}}.
Moreover, plant traits can define species' network roles (e.g.,
specialists vs
generalists)\textsuperscript{\protect\hyperlink{ref-tur2013}{20},\protect\hyperlink{ref-lazaro2013}{37}}.
For example, plant species that occupy reproductive trait space extremes
are more likely to exhibit higher levels of specialisation and be more
reliant on the trait-matching with
pollinators\textsuperscript{\protect\hyperlink{ref-junker2013}{38},\protect\hyperlink{ref-coux2016}{39}}.
Morphological matching between plant and floral visitors often
determines plant-pollinator interactions, and can thus strongly
influence interaction network
structure\textsuperscript{\protect\hyperlink{ref-stang2009}{16},\protect\hyperlink{ref-ibanez2012}{40}}.
Remarkably, the combination of traits have shown to increase the
predictive power of the network
interactions\textsuperscript{\protect\hyperlink{ref-eklof2013}{41}}.
Therefore, considering the different plant reproductive trade-offs which
represent the species strategy within the
network\textsuperscript{\protect\hyperlink{ref-dehling2016}{14}} could
progress our understanding of plant-pollinator interactions. Further, we
know little if the different network patterns generally studied at the
community level are maintained when we scale up at macroecological
scale.

Here, we aim to explore the potential trade-offs among reproductive
traits and how these influence plant-pollinator interactions. First, we
identify the major axes of reproductive trait variation and trade-offs
that determine plant form and function. Second, we investigate how plant
species' position in trait-space interactions with floral visitors.
Finally, we investigate how both the main axes of trait variation, and
individual traits, influence plant species' roles within networks using
a set of complementary interaction network metrics (i.e., interaction
strength, normalized degree and specialization).

\section{RESULTS}\label{results}

\textbf{Plant strategies.} The phylogenetically informed principal
component analysis (pPCA) captured by the first two and three axes
51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and
Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\))
of 0.76. The first principal component (PC1) represented 26.72\% of the
trait variation and indicated a trade-off between flower number and
flower size. We refer to this axis as the `flower number - flower size
trade-off', as already described in previous
studies\textsuperscript{\protect\hyperlink{ref-sargent2007}{42},\protect\hyperlink{ref-kettle2011}{43}}.
Hence, one end of the spectrum comprised species with high investment in
flower number and plant height but small flower size, short style length
and low ovule number. The other end of this spectrum comprised species
that were short in height and invested in large flowers, long styles,
many ovules, but few flowers. The main contributing traits to PC1 were
plant height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Supplementary Table S3) but
style length also contributed moderately on PC1 (loading = -0.33). The
second principal component (PC2) represented 25.05\% of the trait
variation and indicated a trade-off between low and high pollinator
dependence. We refer to this axis as the `pollinator dependence
trade-off'. The main driver of trait variation on PC2 was autonomous
selfing (loading = 0.85) but the other traits (except ovule number) also
made moderate contributions (loadings from 0.27 to 0.4; Supplementary
Table S3). We found that high pollinator dependence was associated with
larger and a higher number of flowers, greater plant height and longer
styles. In contrast, species with high levels of autonomous selfing
tended to have fewer and smaller flowers, had shorter styles and were
shorter in height. Further, PC3 explained a considerable amount of trait
variability (19.17\%) and the main contributors to this axis were style
length (loading = -0.66) and the degree of autonomous selfing (loading =
-0.51). The remaining traits, apart from ovule number, were moderately
correlated to changes on PC3 (loadings from -0.23 to -0.46;
Supplementary Table S3). Thus, because style length was correlated with
all traits on PC3 and was the main driver of trait variation, we refer
to this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA
explained similar variance with the first two principal components
(45.07\%) and similar associations of traits despite some variability in
the loadings (Supplementary Table S4).

\begin{figure}[H]

{\centering \includegraphics{output/figures/unnamed-chunk-1-1}

}

\caption{\textbf{Fig. 1 | Plant life-history strategies.} Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator network studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the dashed lines show the opposed direction of trait variation. The icons at both ends of arrows and dashed lines illustrate the extreme form of the trait continuum.}\label{fig:unnamed-chunk-1}
\end{figure}

We found that most categorical traits were statistically associated with
the first two axes of trait variation (Fig. 2 and Supplementary Table
S2). Flower symmetry, which was only associated with PC2 (Sum of squares
= 8.51, F-value = 14.72, \emph{P} \textless{} 0.01 ), and nectar
provision, which was independent of PC1 and PC2 (PC1: Sum of squares =
0.37, F-value = 0.29 , \emph{P} = 0.59; PC2: Sum of squares = 0.83,
F-value = 1.43, \emph{P} = 0.23) showed lack of statistical association.
In addition, we found (with a Tukey test) statistical differences
between the different levels of categorical traits in the trait space
(Supplementary Fig. S7). Regarding self compatibility, we found larger
differences on PC2 (i.e., species with unisexual flowers that were self
incompatible were statistically differentiated from species with partial
or full self compatibility; Supplementary Fig. S7a and Fig. S7b). Life
forms differed statistically across both axes of trait variation and
followed a gradient of larger life forms (trees and shrubs) with higher
pollinator dependence to smaller ones (herbs) with lower pollinator
dependence (Supplementary Fig. S7c and Fig. S7d). Consequently, lifespan
also followed this gradient but perennial and short lived species only
differed statistically on PC2 (Supplementary Fig. S7e and Fig. S7f).
Species with unisexual flowers (monoecious and dioecious) were clustered
on both extremes of the first two principal components and had the
highest pollinator dependence and highest number of flowers
(Supplementary Fig. S7g and Fig. S7h). Moreover, we found that the
campanulate and capitulum flower shapes were differentiated from tube,
papilionaceous, open and brush shapes in the trait space. The former
morphologies had larger flowers and greater pollinator dependence, while
the latter had higher flower number and greater autonomous selfing
(Supplementary Fig. S7i and Fig. S7j). Regarding flower symmetry,
zygomorphic flowers were associated with lower levels of pollinator
dependence, whereas actinomorphic flowers had higher levels of
pollinator dependence (Supplementary Fig. S7k and Fig. S7l).

\blandscape

\vspace*{-28mm}

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\textbf{Fig. 2 \textbar{} Location of
the different qualitative traits on the trait space.} The panel is
composed by the traits that showed statistical association with the
first two axes of trait variation: compatibility system (a), life form
(b), lifespan (c), breeding system (d), flower shape (e) and flower
symmetry (f).}
\end{figure}

\elandscape

\vspace{5mm}

\textbf{Phylogenetic signal of traits.} We found a strong phylogenetic
signal (\emph{P} \textless{} 0.01) in all quantitative traits
(Supplementary Table S5). The traits that showed the highest
phylogenetic signal were ovule number (\(\lambda = 1\)), pollen grains
per flower (\(\lambda = 1\)) and plant height (\(\lambda = 0.96\)),
followed by flower length (\(\lambda = 0.75\)), flower width
(\(\lambda = 0.73\)), number of flowers per plant (\(\lambda = 0.69\))
and nectar concentration (\(\lambda = 0.65\)). The traits that showed a
moderate phylogenetic signal were inflorescence width
(\(\lambda = 0.57\)), style length (\(\lambda = 0.49\)) and autonomous
selfing (\(\lambda = 0.34\)). Finally, microliters of nectar per flower
showed the lowest phylogenetic signal of all traits
(\(\lambda = 0.14\)).

\textbf{Visitation patterns.} The main axes of trait variation explained
partly presence-absence interactions between plant and floral visitors
(\(conditional R2 = 0.26\); \(marginal R2 = 0.20\)) but little of the
overall visitation rates (\(conditional R2 = 0.31\);
\(marginal R2 = 0.06\)). However, we found relevant trends across the
different floral visitor guilds on both presence-absence and visitation
interactions (Fig. 3). On the pollinator dependence trade-off, all
floral visitor guilds interacted more frequently with plant species with
higher pollinator dependence (PC2; Fig. 3b and Fig. 3e). For
presence-absence interactions we found that all Diptera, Coleoptera and
non-bee Hymenoptera guilds interacted more frequently with plants with
high flower number and small flowers (flower number - flower size
trade-off, PC1; Fig. 3a) but bees and Lepidoptera interacted slightly
more frequently with plant species with low flower number but large
flowers. For presence-absence interactions on PC3 (style length
trade-off; Fig. 3c), we found that bees interacted clearly more with
plant species with long styles and high selfing and the rest of the
guilds interacted slightly more with plant species with short styles and
low selfing. In addition, all guilds other than Syrphids and Lepidoptera
(i.e., all Hymenoptera, non-Syrphidae-Diptera and Coleoptera) showed
greater visitation rates on species with small numerous flowers (PC1;
Fig. 3d). On the style length trade-off, bees, Lepidoptera and
Non-Anthophila-Hymenoptera showed greater visitation rates on plant
species with larger styles and higher levels of selfing; while
Syrphidae, Non-Syrphidae-Diptera and Coleoptera showed higher visitation
rates on species with shorter styles and lower selfing (Fig. 3f).

The additional model for both presence-absence (\(marginal R2 = 0.29\);
\(conditional R2 = 0.19\)) and visitation rate (\(marginal R2 = 0.30\);
\(conditional R2 = 0.03\)) for the most represented families of bees
showed that the family Apidae was the main driver of the observed
patterns and that the contrasting differences between presence-absence
and visitation rate for bees on PC1 (Fig. 3a and Fig. 3d) were driven by
the family Andrenidae which interacted more frequently on
presence-absence interactions with plant species with low number of
flower but large (Supplementary Fig. S8).

\blandscape

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\textbf{Fig. 3 \textbar{} Interaction
(yes/no) and visitation rates across the three main axes of trait
variation per floral visitor guild.} Fitted posterior estimates of the
presence/absence of interaction (a, b and c) and number of visits (d, e
and f) made by the different floral visitors guilds in relation to PC1,
PC2 and PC3. PC1 represents the flower number - flower size trade-off,
PC2 represents the pollinator dependence trade-off and PC3, the style
length trade-off. For visualization purposes, due to large differences
between the visitation rates of bees and the rest of guilds, the number
of visits was log-transformed (Y-axis of lower panel).}
\end{figure}

\elandscape

\textbf{Plant species functional roles.} The variance of the different
plant species-level network metrics was poorly explained by the three
main axes of trait variation (Supplementary Fig. S9; interaction
frequency \textasciitilde{} PCs, \(conditional R2 = 0.11\),
\(marginal R2 = 0.02\); normalized degree \textasciitilde{} PCs,
\(conditional R2 = 0.24\), \(marginal R2 = 0.02\); and, specialization
\textasciitilde{} PCs, \(conditional R2 = 0.37\),
\(marginal R2 = 0.03\)). Overall, the most notable trends were found on
PC1 and PC3 for interaction frequency and specialization. On the flower
number - flower size trade-off (PC1), interaction frequency was higher
for plant species with more flowers but was lower for plant species with
larger flowers. On PC1, specialization showed the opposite trend. On the
style length trade-off (PC3), interaction frequency was lower for plants
with shorter styles and lower autonomous selfing and higher for species
with longer styles and higher autonomous selfing. Again, specialization
showed the opposite trend to interaction frequency.

When we further investigate which combination of traits drive plant
network roles, we show that the regression tree for visitation frequency
was best explained by plant height, nectar concentration and style
length (Fig. 4a). Specifically, species taller than 3.9 m had the
highest interaction frequency, while species that were shorter than 3.9
m and had a nectar concentration lower than 16\% had the lowest
interaction frequency. Normalized degree was best explained by nectar
concentration, pollen grains per flower, plant height, flower width and
autonomous selfing (Fig. 4b). Species with a nectar concentration over
49\% had the highest levels of normalized degree, whereas species with
nectar concentration lower than 49\%, more than 21,000 pollen grains per
flower and height less than 0.78 m had the lowest normalized degree.
Finally, specialization was best explained by plant height, ovule
number, pollen grains per flower and autonomous selfing (Fig. 4c).
Overall, plant species with the highest specialization were shorter than
1.3 m, had more than 14,000 pollen grains per flower and autonomously
self-pollinated less than 11\% of their fruits. In contrast, species
taller or equal than 5.1 m and with lower than 14 ovules per flower had
the lowest specialization values.

\blandscape

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\textbf{Fig. 4 \textbar{} Contribution
of traits in plant's network roles.} Regression tree analysis of
interaction frequency (log-transformed), normalized degree and
specialization for the subset of species with quantitative data for
pollen and nectar traits. The superior value inside the node indicates
the mean value of the different species-level metric and the lower
value, the percentage of species that are considered in each node. Thus,
the top node has the mean value of the named trait for the 100\% of
species. Each node has a yes/no question and when the condition is
fulfilled, the branch turns to the `yes' direction and when not, to the
`no' direction. This rationale is followed in all the regression trees
as indicated in the first branch division of the topmost node of each
tree.}
\end{figure}

\elandscape

\section{DISCUSSION}\label{discussion}

Here, we show that plant species exhibit clear trade-offs among their
vegetative and reproductive traits that determined different interaction
levels with floral visitor guilds. These trade-offs are differentiated
on three main axes of trait variation: (i) flower number - flower size,
(ii) pollinator dependence and (iii) style length. These reproductive
trade-offs somewhat explained the presence or absence of interactions
with floral visitors, but not differences in visitation rate. However,
on both qualitative and quantitative interactions we found marked
associations of floral visitor guilds with the main axes of trait
variation. Further, we found that plant height and floral reward related
traits were best at characterising plant species functional roles within
pollination networks.

Over half of all plant trait variation was captured by the flower number
- flower size and pollinator dependence trade-offs. Trait variation on
these two axes was associated with the `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{44}}
life-history strategies, as indicated by the different floral and
reproductive biology traits associated with plant height, life form and
lifespan. The `slow' part of this continuum (i.e., tall trees and
shrubs) included plant species with many flowers, few ovules, higher
pollinator dependence, frequent occurrence of self-incompatibility and
more complex breeding systems (e.g., monoecious and dioecious species).
In contrast, plant species that employed the `fast' strategy (i.e.,
short herbs), had fewer flowers, more ovules, frequent occurrence of
self-compatibility and lower pollinator dependence. Further, on the
first two axes of trait variation, we found additional support for the
previously described positive association between higher outcrossing
rate and larger floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{31}}. The
positive correlation between larger floral display and higher pollinator
dependence in our dataset further confirmed this trend (see
Supplementary Fig. S10).

Despite the low predictive power of the main trait variation axes for
broad-level interaction patterns (presence-absence of interactions and
visitation rate), we found changes in the interaction patterns among and
within floral visitor guilds across these axes that suggest plant
life-history strategies strongly influence plant-pollinator
interactions. For example, all floral visitor guilds visited plant
species with higher pollinator dependence more frequently, and high
pollinator dependence was associated with large floral displays and
greater pollen quantities (Fig. 1 and Supplementary Fig. S6). This trend
is consistent with previous studies that show plant species with higher
reproductive investment tend to be visited by pollinators more
frequently\textsuperscript{\protect\hyperlink{ref-lazaro2013}{37},\protect\hyperlink{ref-hegland2005}{45},\protect\hyperlink{ref-kaiser2014}{46}}.
In regard to the flower number - flower size and style length
trade-offs, different pollinator guilds showed contrasting visitation
rates across the continuum of trait variation, which could be associated
with different pollination syndromes at a macroecological scale. For
instance, bees and syrphid flies were clearly associated with opposing
life-strategies on PC1 and PC3 (Fig. 3) suggesting possible niche
partitioning\textsuperscript{\protect\hyperlink{ref-palmer2003}{47},\protect\hyperlink{ref-phillips2020}{48}}
between these two guilds. However, despite floral rewards, related
traits were among the best at characterising species functional roles
(Fig. 4), yet the association between reproductive trade-offs and floral
visitors did not account for floral rewards because there was
insufficient data available. More detailed exploration of reproductive
trade-offs in conjunction with floral rewards is needed to help
elucidate plant-pollinator associations. In any case, it is worth noting
that other local factors such as species relative abundances, surely
explain part of the observed
variability\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{17},\protect\hyperlink{ref-vazquez2007}{49},\protect\hyperlink{ref-encinas2012}{50}}
that reproductive trade-offs do not.

To conclude, we provide the first description of plant reproductive
trade-offs using a large global dataset of plant traits. We identified
the major reproductive strategies of flowering plants and how these
strategies influence interactions with different floral visitor guilds.
Although the explained variation that we found in the first two axes is
lower than previous studies of vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{24},\protect\hyperlink{ref-carmona2021}{25}}
it is consistent with the largest and most recent study that has
characterised plant life strategies with vegetative and reproductive
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{12}}. Future
work needs to integrate the reproductive compromises that we have
identified with vegetative and physiological trade-offs to create a more
comprehensive spectrum of plant trait variation. Further, the varying
level of phylogenetic signal among traits deserves further attention to
understand evolutionary changes on mating and flower morphology in
response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{51},\protect\hyperlink{ref-mackin2021}{52}}.
Finally, including plant-pollinator networks from unrepresented areas of
the world and a more complete description of plant reproductive
trade-offs is essential for a better understanding of the global
patterns in plant-pollinator interactions.

\section{MATERIALS AND METHODS}\label{materials-and-methods}

\textbf{Plant-pollinator network studies.} We selected 28 studies from
18 different countries that constituted a total of 64 plant-pollinator
networks. These studies recorded plant-pollinator interactions in
natural systems and were selected so that we had broad geographical
representation. Although these studies differ in sampling effort and
methodology, all studies provided information about plant-pollinator
interactions (weighted and non-weighted), which we used to build a
database of plant species that are likely to be animal pollinated. Many
of these networks are freely available either as published
studies\textsuperscript{\protect\hyperlink{ref-fortuna2010}{53},\protect\hyperlink{ref-carvalheiro2014}{54}}
or available in online archives (e.g., `The Web of
Life'\textsuperscript{\protect\hyperlink{ref-fortuna2010}{53}} and
`Mangal'\textsuperscript{\protect\hyperlink{ref-poisot2016}{55}}. In
total, our network dataset (see Supplementary Table S1) constituted 60
weighted (interaction frequency) and 4 unweighted (presence/absence of
the interaction) networks, each sampled at a unique location and year,
as well as eight meta-webs where interactions were pooled across several
locations and multiple years.

\textbf{Taxonomy of plants and pollinators.} All species names, genera,
families and orders were retrieved and standardized from the taxonomy
data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for
plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the
R package
\emph{taxize}\textsuperscript{\protect\hyperlink{ref-chamberlain2020}{56}}.
We filled the `not found' searches manually using
\url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for
plants and \url{http://www.catalogueoflife.org/} for floral visitors.

\textbf{Functional traits.} We selected 20 different functional traits
based on their relevance to plant reproduction and data availability
(Table 1). These included twelve quantitative and eight categorical
traits belonging to three broader trait groupings (13 floral, 4
reproductive biology and 3 vegetative, Supplementary Information). For
each plant species, we undertook an extensive literature and online
search across a wide range of resources (plant databases, online floras,
books, journals and images). From a total of 30,120 cells (20 columns ×
1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset,
see Supplementary Fig. S1 for missing values information for each
trait).

\textbf{Phylogenetic Distance.} We calculated the phylogenetic distance
between different plant species using the function \emph{get\_tree} from
the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}),
which downloads phylogenetic distances from the extended R
implementation of the Open Tree of
Life\textsuperscript{\protect\hyperlink{ref-smith2018}{57},\protect\hyperlink{ref-jin2019}{58}}.

\singlespacing

\begin{table}

\caption{\label{tab:unnamed-chunk-5}\textbf{Table 1 | Quantitative and categorical traits used in this study.}}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{}cl>{}ccl}
\toprule
\multicolumn{2}{c}{\textbf{Quantitative traits}} & \multicolumn{3}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Type} & \textbf{Traits} & \textbf{Type} & \textbf{Traits} & \textbf{Categories}\\
\midrule
\textbf{Vegetative} & Plant height (m) & \textbf{Vegetative} & Lifepan & \makecell[l]{Short-lived \\ Perennial}\\
\addlinespace
\textbf{Floral} & Flower width (mm) & \textbf{Vegetative} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree}\\
\addlinespace
\textbf{Floral} & Flower length (mm) & \textbf{Floral} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube}\\
\addlinespace
\textbf{Floral} & Inflorescence width (mm) & \textbf{Floral} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic}\\
\addlinespace
\textbf{Floral} & Style length (mm) & \textbf{Floral} & Nectar & \makecell[l]{Presence \\ Absence}\\
\addlinespace
\textbf{Floral} & Ovules per flower & \textbf{\makecell[c]{Reproductive \\ biology}} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High}\\
\addlinespace
\textbf{Floral} & Flowers per plant & \textbf{\makecell[c]{Reproductive \\ biology}} & Compatibility system & \makecell[l]{Self-incomp. \\ Part. self-comp. \\ Self-comp.}\\
\addlinespace
\textbf{Floral} & Nectar ($\mu$l) & \textbf{\makecell[c]{Reproductive \\ biology}} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious}\\
\addlinespace
\textbf{Floral} & Nectar (mg) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Nectar concentration ($\%$) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Pollen grains per flower & \textbf{} &  & \\
\addlinespace
\textbf{\makecell[c]{Reproductive \\ biology}} & Autonomous selfing (fruit set) & \textbf{} &  & \\
\bottomrule
\end{tabular}
\end{table}

\doublespacing

\textbf{Data Imputation.} Trait missing values were imputed with the
function
\emph{missForest}\textsuperscript{\protect\hyperlink{ref-stekhoven2012}{59}}
which allows imputation of data sets with continuous and categorical
variables. We accounted for the phylogenetic distance among species on
the imputation process by including the eigenvectors of a principal
component analysis of the phylogenetic distance (PCoA) which has been
shown to improve the performance of
\emph{missForest}\textsuperscript{\protect\hyperlink{ref-penone2014}{60}}.
To extract the eigenvectors, we used the function \emph{PVRdecomp} from
the package
\emph{PVR}\textsuperscript{\protect\hyperlink{ref-santos2018}{61}} based
on a previous conceptual framework that considers phylogenetic
eigenvectors\textsuperscript{\protect\hyperlink{ref-diniz-filho2012}{62}}.
Although the variable of autonomous selfing had a high percentage of
missing values (68\%), we were able to solve this by back transforming
the qualitative column of autonomous selfing to numerical. The
categories of `none', `low', `medium' and `high' were converted to
representative percentages of each category 0\%, 13\%, 50.5\% and 88\%
respectively. This reduced the percentage of missing values for this
column from 68\% to 35\% and allowed the imputation of this variable.
However, we were unable to include nectar and pollen traits on the
imputation process because of the high percentage of missing values
(Supplementary Fig. S1). Hence, the imputed dataset had 1,506 species,
seven categorical and eight numerical variables and 5.79\% of missing
values. Further, we conducted an additional imputation on the filtered
raw data by filled cells of either pollen grains per flower or
microlitres of nectar. This subset comprised 755 species, 8.01\% missing
values and all traits but milligrams of nectar (\textasciitilde{}50\% of
missing values) were included in the imputation process.

\textbf{Plant strategies.} We explored the trade-offs between different
quantitative plant functional traits with a phylogenetically informed
Principal Component Analysis (pPCA). We did not include the quantitative
variables of flower length and inflorescence width because they were
highly and moderately correlated to flower width respectively (Pearson's
correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation
= 0.36, \emph{P} \textless{} 0.01 respectively), and thus we avoided
overemphasizing flower size on the spectrum of trait variation. Although
qualitative traits were not included in the dimensionality reduction
analysis, we also investigated the association of the different
qualitative traits with the main axes of trait variation. Prior to the
analyses, we excluded outliers and standardized the data. Due to the
high sensitivity of dimensionality reduction to outliers, we excluded
values within the 2.5th--97.5th percentile
range\textsuperscript{\protect\hyperlink{ref-legendre2012}{63}}, and
thus our final dataset had 1,236 species. Then, we log transformed the
variables to reduce the influence of outliers and z-transformed (X= 0,
SD=1) so that all variables were within the same numerical range. We
performed the pPCA using the function \emph{phyl.pca} from the package
\emph{phytools}\textsuperscript{\protect\hyperlink{ref-revell2012}{64}}
with the method lambda (\(\lambda\)) that calculates the phylogenetic
correlation between 0 (phylogenetic independence) and 1 (shared
evolutionary history) and we implemented the mode covariance because
values for each variables were on the same scale following
transformation\textsuperscript{\protect\hyperlink{ref-abdi2010}{65}}.
Moreover, to corroborate that our imputation of missing values did not
affect our results, we conducted a pPCA on the full dataset without
missing values (see Supplementary Fig. S2). We found little difference
between the explained variance with the imputed dataset (51.08\%) and
the dataset without missing values (52.87\%). In addition, the loadings
on each principal component had a similar contribution and correlation
patterns, with the exception of plant height which showed slight
variations between the imputed and non-imputed dataset. Finally, we
conducted an additional phylogenetic informed principal component
analysis for the subset of species with pollen and nectar quantity. For
this, we included all quantitative traits considered in the main pPCA
plus pollen grains and microlitres of nectar per flower.

\textbf{Phylogenetic signal of traits.} We calculated the phylogenetic
signal of the different quantitative traits on the imputed dataset with
the full set of species (N = 1,506) with the package \emph{phytools}
version
\emph{0.7-70}\textsuperscript{\protect\hyperlink{ref-revell2012}{64}}
and we used Pagel's \(\lambda\) as a measurement of the phylogenetic
signal. However, for pollen and nectar traits, phylogenetic signal was
calculated only on the subset of species that had quantitative
information for these traits (N = 755).

\textbf{Network analyses.} Analyses were conducted on the subset of 60
weighted networks sampled in a unique flowering season and site, which
included 556 plant and 1,126 pollinator species. These networks were
analysed in their qualitative (presence-absence) and quantitative
(interaction frequency) form. First, we analysed the binary version of
these weighted networks with presence-absence information that assumes
equal weight across interactions. Second, we analysed the untransformed
weighted networks with interaction frequency that accounts for the
intensity of the interaction. Although floral visitors are not always
pollinators and interaction frequency does not consider each pollinator
species
efficiency\textsuperscript{\protect\hyperlink{ref-ballantyne2015}{66}},
interaction frequency can provide valuable information of the
contribution of floral visitors to
pollination\textsuperscript{\protect\hyperlink{ref-vazquez2005}{67},\protect\hyperlink{ref-vazquez2012}{68}}.
In total, our network dataset (excluding meta-webs and non-weighted
networks) included 2,256 interactions of Anthophila-Hymenoptera (i.e.,
bees) with plants, 1,768 non-Syrphidae-Diptera interactions, 845
Syrphidae interactions, 437 Lepidoptera interactions, 432 Coleoptera
interactions and 362 non-Anthophila-Hymenoptera interactions. Sampling
methods varied across networks but this was accounted for in analyses by
considering them in the random effects of the modelling process. All
analyses were conducted in R version \emph{4.0.3}.

\textbf{Visitation patterns.} We used Bayesian modelling (see below for
details) to explore the effect of floral visitor groups and the main
axes of trait variation (pPCA with imputed dataset) on both qualitative
(presence/absence) and quantitative (visitation rate) floral
interactions per plant species. For this, we divided floral visitors
into six main guilds that differ in life form, behaviour and are likely
to play a similar ecological role: (i) Hymenoptera-Anthophila (bees),
(ii) Hymenoptera-non-Anthophila (non-bee Hymenoptera), (iii)
Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi)
Coleoptera. Moreover, because Hymenoptera-Anthophila was the most
represented group with 2,256 records and had the highest frequency of
visits of all groups, we also explored the presence-absence of
interaction and visitation rate of the main Hymenoptera-Anthophila
families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae)
on the trait space. In addition, we found that \emph{Apis mellifera} was
the floral visitor with the largest proportion of records counted
(7.55\% of the total). This finding is consistent with previous research
showing that \emph{A. mellifera} was the most frequent floral visitor in
a similar dataset of 80 plant-pollinator networks in natural
ecosystems\textsuperscript{\protect\hyperlink{ref-hung2018}{69}}. Hence,
to control for the effect of \emph{A. mellifera} on the observed
visitation patterns of bees, we conducted an analogous analysis with
presence-absence of interaction and visitation rate excluding \emph{A.
mellifera}. We found that \emph{A. mellifera}, was partly driving some
of the observed trends on PC1 (Supplementary Fig. S3). However, we did
not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R
package
\emph{brms}\textsuperscript{\protect\hyperlink{ref-burkner2017}{70}}. We
modelled the frequency of visits as a function of the main axes of plant
trait variation and their interactions with floral visitor functional
groups (Visits \textasciitilde{} PC1 x FGs + PC2 x FGs + PC3 x FGs).
Because we were interested in possible differences in the visitation
patterns among floral visitors groups to plants with different
strategies, we included interactions between the main axes of trait
variation (PC1, PC2 and PC3) and the floral visitor guilds. In this
model, we added a nested random effect of networks nested within the
study system to capture the variation in networks among studies and
within networks. Moreover, we included the phylogenetic covariance
matrix as a random factor due to the possible shared evolutionary
histories of species and therefore lack of independence across them. We
specified this model with a zero inflated negative binomial distribution
and weakly informative priors from the brms function. We run this model
for 3,000 iterations and with previous 1,000 warm up iterations. We set
delta (\(\Delta\)) to 0.99 to avoid divergent transitions and visualized
the posterior predictive checks with the function \emph{pp\_check} using
the \emph{bayesplot}
package\textsuperscript{\protect\hyperlink{ref-gabry2019}{71}}.

\textbf{Plant species functional roles.} We investigated whether
different quantitative traits determined plant species functional roles
using Bayesian modelling and regression trees. For this, we selected
simple and complementary species-level network metrics commonly applied
in bipartite network
studies\textsuperscript{\protect\hyperlink{ref-dormann2008}{72}} with a
straightforward ecological interpretation relevant to our research
goals. The different plant species-level metrics were: (i) sum of visits
per plant species; (ii) normalized degree, calculated as the number of
links per plant species divided by the total possible number of
partners; and (iii)
specialization\textsuperscript{\protect\hyperlink{ref-bluthgen2006}{73}},
which measures the deviation of an expected random choice of the
available interaction partners and ranges between 0 (maximum
generalization) and 1 (maximum specialization). Normalized degree and
specialization were calculated with the \emph{specieslevel} function
from the R package
\emph{bipartite}\textsuperscript{\protect\hyperlink{ref-dormann2008}{72}}.

First, we modelled the distinct plant species metrics (sum of visits,
normalized degree and plant specialization) as a function of the three
main axes of trait variation (plant species level metric
\textasciitilde{} PC1 + PC2 + PC3). For each response variable (i.e.,
each plant species level metric), we used different distribution
families (zero inflated negative binomial for the sum of visits, weibull
for normalized degree and zero one inflated beta for specialization).
Finally, we used the same random factors, model settings and conducted
the same posterior predictive checks for each model as detailed above in
the `visitation patterns section'.

Second, to better understand these complex trait relationships, we used
regression trees. Regression trees are recursive algorithms which can
detect complex relationships among predictors and allow identification
of the relevance of specific trait combinations on species functional
roles. We focused exclusively on quantitative traits because almost all
categorical traits were statistically associated with the first two axes
of trait variation (Supplementary Table S2). We conducted this analysis
using the \emph{rpart}
package\textsuperscript{\protect\hyperlink{ref-therneau2015}{74}}
version \emph{4.1-15} with method \emph{`anova'} with a minimum of 50
observations per terminal node and we used the \emph{rpart.plot}
package\textsuperscript{\protect\hyperlink{ref-milborrow2015}{75}}
version \emph{3.0.9} to plot the regression trees. We considered the
species level indices as response variables (interaction frequency,
normalized degree and specialization) and we performed one regression
tree per metric using the different quantitative traits as predictors.
We calculated two regression trees per plant species-level metric, one
for the full set of species and another for the subset of species for
which we had pollen and nectar traits. We focused on regression trees
that included floral rewards because they consistently showed pollen and
nectar traits as being the best for explaining the different
species-level metrics (see Supplementary Fig. S4).

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-barrett2002}{}
1. Barrett, S. C. H. The evolution of plant sexual diversity. \emph{Nat.
Rev. Genet.} \textbf{3}, 274--284 (2002).

\hypertarget{ref-schiestl2013}{}
2. Schiestl, F. P. \& Johnson, S. D. Pollinator-mediated evolution of
floral signals. \emph{Trends Ecol. Evol.} \textbf{28}, 307--315 (2013).

\hypertarget{ref-fenster2004}{}
3. Fenster, C. B., Armbruster, W. S., Wilson, P., Dudash, M. R. \&
Thomson, J. D. Pollination Syndromes and Floral Specialization.
\emph{Annu. Rev. Ecol. Evol. Syst.} \textbf{35}, 375--403 (2004).

\hypertarget{ref-dellinger2020}{}
4. Dellinger, A. S. Pollination syndromes in the 21st century: Where do
we stand and where may we go? \emph{New Phytol.} \textbf{228},
1193--1213 (2020).

\hypertarget{ref-roddy2021}{}
5. Roddy, A. B. \emph{et al.} Towards the flower economics spectrum.
\emph{New Phytol.} \textbf{229}, 665--672 (2021).

\hypertarget{ref-baude2016}{}
6. Baude, M. \emph{et al.} Historical nectar assessment reveals the fall
and rise of floral resources in britain. \emph{Nature} \textbf{530},
85--88 (2016).

\hypertarget{ref-munoz2016}{}
7. Munoz, F., Violle, C. \& Cheptou, P.-O. CSR ecological strategies and
plant mating systems: Outcrossing increases with competitiveness but
stress-tolerance is related to mixed mating. \emph{Oikos} \textbf{125},
1296--1303 (2016).

\hypertarget{ref-grossenbacher2017}{}
8. Grossenbacher, D. L. \emph{et al.} Self-compatibility is
over-represented on islands. \emph{New Phytol.} \textbf{215}, 469--478
(2017).

\hypertarget{ref-moeller2017}{}
9. Moeller, D. A. \emph{et al.} Global biogeography of mating system
variation in seed plants. \emph{Ecol. Lett.} \textbf{20}, 375--384
(2017).

\hypertarget{ref-sargent2008}{}
10. Sargent, R. D. \& Ackerly, D. D. Plant--pollinator interactions and
the assembly of plant communities. \emph{Trends Ecol. Evol.}
\textbf{23}, 123--130 (2008).

\hypertarget{ref-rech2016}{}
11. Rech, A. R. \emph{et al.} The macroecology of animal versus wind
pollination: Ecological factors are more important than historical
climate stability. \emph{Plant Ecol. Divers.} \textbf{9}, 253--262
(2016).

\hypertarget{ref-salguero2016}{}
12. Salguero-Gómez, R. \emph{et al.} Fast-slow continuum and
reproductive strategies structure plant life-history variation
worldwide. \emph{Proc. Natl. Acad. Sci. U.S.A.} \textbf{113}, 230--235
(2016).

\hypertarget{ref-ruger2018}{}
13. Rüger, N. \emph{et al.} Beyond the fastSlow continuum: Demographic
dimensions structuring a tropical tree community. \emph{Ecol. Lett.}
\textbf{21}, 1075--1084 (2018).

\hypertarget{ref-dehling2016}{}
14. Dehling, D. M., Jordano, P., Schaefer, H. M., Böhning-Gaese, K. \&
Schleuning, M. Morphology predicts species' functional roles and their
degree of specialization in plant--frugivore interactions. \emph{Proc.
Royal Soc. B} \textbf{283}, 20152444 (2016).

\hypertarget{ref-rosas2014}{}
15. Rosas-Guerrero, V. \emph{et al.} A quantitative review of
pollination syndromes: Do floral traits predict effective pollinators?
\emph{Ecol. Lett.} \textbf{17}, 388--400 (2014).

\hypertarget{ref-stang2009}{}
16. Stang, M., Klinkhamer, P. G. L., Waser, N. M., Stang, I. \& van der
Meijden, E. Size-specific interaction patterns and size matching in a
plantPollinator interaction web. \emph{Ann. Bot.} \textbf{103},
1459--1469 (2009).

\hypertarget{ref-bartomeus2016}{}
17. Bartomeus, I. \emph{et al.} A common framework for identifying
linkage rules across different types of interactions. \emph{Funct.
Ecol.} \textbf{30}, 1894--1903 (2016).

\hypertarget{ref-waser1996}{}
18. Waser, N. M., Chittka, L., Price, M. V., Williams, N. M. \&
Ollerton, J. Generalization in pollination systems, and why it matters.
\emph{Ecology} \textbf{77}, 1043--1060 (1996).

\hypertarget{ref-olesen2002}{}
19. Olesen, J. M. \& Jordano, P. Geographic patterns in
plant--pollinator mutualistic networks. \emph{Ecology} \textbf{83},
2416--2424 (2002).

\hypertarget{ref-tur2013}{}
20. Tur, C., Castro-Urgal, R. \& Traveset, A. Linking Plant
Specialization to Dependence in Interactions for Seed Set in Pollination
Networks. \emph{PLoS One} \textbf{8}, e78294 (2013).

\hypertarget{ref-devaux2014}{}
21. Devaux, C., Lepers, C. \& Porcher, E. Constraints imposed by
pollinator behaviour on the ecology and evolution of plant mating
systems. \emph{J. Evol. Biol.} \textbf{27}, 1413--1430 (2014).

\hypertarget{ref-kattge2011}{}
22. Kattge, J. \emph{et al.} TRY a global database of plant traits.
\emph{Glob. Chang. Biol.} \textbf{17}, 2905--2935 (2011).

\hypertarget{ref-salguero2015}{}
23. Salguero-Gómez, R. \emph{et al.} The compadre Plant Matrix Database:
An open online repository for plant demography. \emph{J. Ecol.}
\textbf{103}, 202--218 (2015).

\hypertarget{ref-diaz2016}{}
24. Díaz, S. \emph{et al.} The global spectrum of plant form and
function. \emph{Nature} \textbf{529}, 167--171 (2016).

\hypertarget{ref-carmona2021}{}
25. Carmona, C. P. \emph{et al.} Erosion of global functional diversity
across the tree of life. \emph{Sci. Adv.} \textbf{7}, eabf2675 (2021).

\hypertarget{ref-wright2004}{}
26. Wright, I. J. \emph{et al.} The worldwide leaf economics spectrum.
\emph{Nature} \textbf{428}, 821--827 (2004).

\hypertarget{ref-chave2009}{}
27. Chave, J. \emph{et al.} Towards a worldwide wood economics spectrum.
\emph{Ecol. Lett.} \textbf{12}, 351--366 (2009).

\hypertarget{ref-laughlin2021}{}
28. Laughlin, D. C. \emph{et al.} Root traits explain plant species
distributions along climatic gradients yet challenge the nature of
ecological trade-offs. \emph{Nat. Ecol. Evol.} 1--12 (2021).

\hypertarget{ref-evojtko2020}{}
29. E-Vojtkó, A., Bello, F. de, Durka, W., Kühn, I. \& Götzenberger, L.
The neglected importance of floral traits in trait-based plant community
assembly. \emph{J. Veg. Sci.} \textbf{31}, 529--539 (2020).

\hypertarget{ref-barrett2003}{}
30. Barrett, S. C. H. Mating strategies in flowering plants: The
outcrossing-selfing paradigm and beyond. \emph{Philos. Trans. R. Soc.
Lond. B Biol. Sci.} \textbf{358}, 991--1004 (2003).

\hypertarget{ref-goodwillie2010}{}
31. Goodwillie, C. \emph{et al.} Correlated evolution of mating system
and floral display traits in flowering plants and its implications for
the distribution of mating system variation. \emph{New Phytol.}
\textbf{185}, 311--321 (2010).

\hypertarget{ref-ollerton2011}{}
32. Ollerton, J., Winfree, R. \& Tarrant, S. How many flowering plants
are pollinated by animals? \emph{Oikos} \textbf{120}, 321--326 (2011).

\hypertarget{ref-rodger2021}{}
33. Rodger, J. G. \emph{et al.} Widespread vulnerability of flowering
plant seed production to pollinator declines. \emph{Science Advances}
\textbf{7}, eabd3524 (2021).

\hypertarget{ref-bartomeus2013}{}
34. Bartomeus, I. Understanding Linkage Rules in Plant-Pollinator
Networks by Using Hierarchical Models That Incorporate Pollinator
Detectability and Plant Traits. \emph{PLoS One} \textbf{8}, e69200
(2013).

\hypertarget{ref-olito2015}{}
35. Olito, C. \& Fox, J. W. Species traits and abundances predict
metrics of plantPollinator network structure, but not pairwise
interactions. \emph{Oikos} \textbf{124}, 428--436 (2015).

\hypertarget{ref-rowe2020}{}
36. Rowe, L. \emph{et al.} Flower traits associated with the visitation
patterns of bees. \emph{Oecologia} \textbf{193}, 511--522 (2020).

\hypertarget{ref-lazaro2013}{}
37. Lázaro, A., Jakobsson, A. \& Totland, Ø. How do pollinator
visitation rate and seed set relate to species' floral traits and
community context? \emph{Oecologia} \textbf{173}, 881--893 (2013).

\hypertarget{ref-junker2013}{}
38. Junker, R. R. \emph{et al.} Specialization on traits as basis for
the niche-breadth of flower visitors and as structuring mechanism of
ecological networks. \emph{Funct. Ecol.} \textbf{27}, 329--341 (2013).

\hypertarget{ref-coux2016}{}
39. Coux, C., Rader, R., Bartomeus, I. \& Tylianakis, J. M. Linking
species functional roles to their network roles. \emph{Ecol. Lett.}
\textbf{19}, 762--770 (2016).

\hypertarget{ref-ibanez2012}{}
40. Ibanez, S. Optimizing size thresholds in a plant-pollinator
interaction web: Towards a mechanistic understanding of ecological
networks. \emph{Oecologia} \textbf{170}, 233--242 (2012).

\hypertarget{ref-eklof2013}{}
41. Eklöf, A. \emph{et al.} The dimensionality of ecological networks.
\emph{Ecol. Lett.} \textbf{16}, 577--583 (2013).

\hypertarget{ref-sargent2007}{}
42. Sargent, R. D., Goodwillie, C., Kalisz, S. \& Ree, R. H.
Phylogenetic evidence for a flower size and number trade-off. \emph{Am.
J. Bot.} \textbf{94}, 2059--2062 (2007).

\hypertarget{ref-kettle2011}{}
43. Kettle, C. J. \emph{et al.} Ecological Implications of a Flower
Size/Number Trade-Off in Tropical Forest Trees. \emph{PLoS One}
\textbf{6}, e16111 (2011).

\hypertarget{ref-healy2019}{}
44. Healy, K., Ezard, T. H. G., Jones, O. R., Salguero-Gómez, R. \&
Buckley, Y. M. Animal life history is shaped by the pace of life and the
distribution of age-specific mortality and reproduction. \emph{Nat.
Ecol. Evol.} \textbf{3}, 1217--1224 (2019).

\hypertarget{ref-hegland2005}{}
45. Hegland, S. J. \& Totland, Ø. Relationships between species' floral
traits and pollinator visitation in a temperate grassland.
\emph{Oecologia} \textbf{145}, 586--594 (2005).

\hypertarget{ref-kaiser2014}{}
46. Kaiser-Bunbury, C. N., Vázquez, D. P., Stang, M. \& Ghazoul, J.
Determinants of the microstructure of plantPollinator networks.
\emph{Ecology} \textbf{95}, 3314--3324 (2014).

\hypertarget{ref-palmer2003}{}
47. Palmer, T. M., Stanton, M. L. \& Young, T. P. Competition and
coexistence: Exploring mechanisms that restrict and maintain diversity
within mutualist guilds. \emph{Am. Nat.} \textbf{162}, S63--S79 (2003).

\hypertarget{ref-phillips2020}{}
48. Phillips, R. D., Peakall, R., van der Niet, T. \& Johnson, S. D.
Niche perspectives on plantPollinator interactions. \emph{Trends Plant
Sci.} \textbf{25}, 779--793 (2020).

\hypertarget{ref-vazquez2007}{}
49. Vázquez, D. P. \emph{et al.} Species abundance and asymmetric
interaction strength in ecological networks. \emph{Oikos} \textbf{116},
1120--1127 (2007).

\hypertarget{ref-encinas2012}{}
50. Encinas-Viso, F., Revilla, T. A. \& Etienne, R. S. Phenology drives
mutualistic network structure and diversity. \emph{Ecol. Lett.}
\textbf{15}, 198--208 (2012).

\hypertarget{ref-gervasi2017}{}
51. Gervasi, D. D. L. \& Schiestl, F. P. Real-time divergent evolution
in plants driven by pollinators. \emph{Nat. Commun.} \textbf{8}, 14691
(2017).

\hypertarget{ref-mackin2021}{}
52. Mackin, C. R., Peña, J. F., Blanco, M. A., Balfour, N. J. \&
Castellanos, M. C. Rapid evolution of a floral trait following
acquisition of novel pollinators. \emph{J. Ecol.} \textbf{109},
2234--2246 (2021).

\hypertarget{ref-fortuna2010}{}
53. Fortuna, M. A. \emph{et al.} Nestedness versus modularity in
ecological networks: Two sides of the same coin? \emph{J. Anim. Ecol.}
\textbf{79}, 811--817 (2010).

\hypertarget{ref-carvalheiro2014}{}
54. Carvalheiro, L. G. \emph{et al.} The potential for indirect effects
between co-flowering plants via shared pollinators depends on resource
abundance, accessibility and relatedness. \emph{Ecol. Lett.}
\textbf{17}, 1389--1399 (2014).

\hypertarget{ref-poisot2016}{}
55. Poisot, T. \emph{et al.} Mangal making ecological network analysis
simple. \emph{Ecography} \textbf{39}, 384--390 (2016).

\hypertarget{ref-chamberlain2020}{}
56. Chamberlain, S. \emph{et al.} Taxize: Taxonomic information from
around the web. r package version 0.9.99 (2020).~at
\href{https://CRAN.R-project.org/package=taxize}{https://cran.R-project.org/package=taxize}.

\hypertarget{ref-smith2018}{}
57. Smith, S. A. \& Brown, J. W. Constructing a broadly inclusive seed
plant phylogeny. \emph{Am. J. Bot.} \textbf{105}, 302--314 (2018).

\hypertarget{ref-jin2019}{}
58. Jin, Y. \& Qian, H. V.PhyloMaker: An R package that can generate
very large phylogenies for vascular plants. \emph{Ecography}
\textbf{42}, 1353--1359 (2019).

\hypertarget{ref-stekhoven2012}{}
59. Stekhoven, D. J. \& Bühlmann, P. MissForestNon-parametric missing
value imputation for mixed-type data. \emph{Bioinformatics} \textbf{28},
112--118 (2012).

\hypertarget{ref-penone2014}{}
60. Penone, C. \emph{et al.} Imputation of missing data in life-history
trait datasets: Which approach performs the best? \emph{Methods Ecol.
Evol.} \textbf{5}, 961--970 (2014).

\hypertarget{ref-santos2018}{}
61. Chamberlain, S. \emph{et al.} Package `PVR'. r package version 0.3
(2018).~at
\href{https://CRAN.R-project.org/package=PVR}{https://cran.R-project.org/package=PVR}.

\hypertarget{ref-diniz-filho2012}{}
62. Diniz-Filho, J. A. F. \emph{et al.} On the selection of phylogenetic
eigenvectors for ecological analyses. \emph{Ecography} \textbf{35},
239--249 (2012).

\hypertarget{ref-legendre2012}{}
63. Legendre, P. \& Legendre, L. \emph{Numerical ecology}. (Elsevier,
Amsterdam, 2012).

\hypertarget{ref-revell2012}{}
64. Revell, L. J. Phytools: An R package for phylogenetic comparative
biology (and other things). \emph{Methods Ecol. Evol.} \textbf{3},
217--223 (2012).

\hypertarget{ref-abdi2010}{}
65. Abdi, H. \& Williams, L. J. Principal component analysis.
\emph{WIREs Comp. Stats.} \textbf{2}, 433--459 (2010).

\hypertarget{ref-ballantyne2015}{}
66. Ballantyne, G., Baldock, K. C. R. \& Willmer, P. G. Constructing
more informative plantPollinator networks: Visitation and pollen
deposition networks in a heathland plant community. \emph{Proc. Royal
Soc. B} \textbf{282}, 20151130 (2015).

\hypertarget{ref-vazquez2005}{}
67. Vázquez, D. P., Morris, W. F. \& Jordano, P. Interaction frequency
as a surrogate for the total effect of animal mutualists on plants.
\emph{Ecol. Lett.} \textbf{8}, 1088--1094 (2005).

\hypertarget{ref-vazquez2012}{}
68. Vázquez, D. P. \emph{et al.} The strength of plantPollinator
interactions. \emph{Ecology} \textbf{93}, 719--725 (2012).

\hypertarget{ref-hung2018}{}
69. Hung, K.-L. J., Kingston, J. M., Albrecht, M., Holway, D. A. \&
Kohn, J. R. The worldwide importance of honey bees as pollinators in
natural habitats. \emph{Proc. Royal Soc. B} \textbf{285}, 20172140
(2018).

\hypertarget{ref-burkner2017}{}
70. Bürkner, P.-C. Brms: An R package for Bayesian multilevel models
using Stan. \emph{J. Stat. Softw.} \textbf{80}, 1--28 (2017).

\hypertarget{ref-gabry2019}{}
71. Gabry, J., Simpson, D., Vehtari, A., Betancourt, M. \& Gelman, A.
Visualization in Bayesian workflow. \emph{J. R. Stat. Soc. Ser. A Stat.
Soc.} \textbf{182}, 389--402 (2019).

\hypertarget{ref-dormann2008}{}
72. Dormann, C. F., Gruber, B. \& Fründ, J. Introducing the bipartite
package: Analysing ecological networks. \emph{interaction} \textbf{1},
(2008).

\hypertarget{ref-bluthgen2006}{}
73. Blüthgen, N., Menzel, F. \& Blüthgen, N. Measuring specialization in
species interaction networks. \emph{BMC Ecol.} \textbf{6}, 9 (2006).

\hypertarget{ref-therneau2015}{}
74. Therneau, T., Atkinson, B., Ripley, B. \& Ripley, M. B. Package
`rpart'. r package version 4.1-15 (2015).~at
\href{https://CRAN.R-project.org/package=rpart}{https://cran.R-project.org/package=rpart}.

\hypertarget{ref-milborrow2015}{}
75. Milborrow, S. Rpart. plot: Plot rpart Models. An Enhanced Version of
plot. rpart. r package version 3.0.9 (2015).~at
\href{https://CRAN.R-project.org/package=rpart.plot}{https://cran.R-project.org/package=rpart.plot}.

\section{Acknowledgements}\label{acknowledgements}

We thank all researchers that made their data available for our
analysis. We thank Bryony Wilcox, Greg Bible, Mercedes Sanchez-Lanuza
and David Ragel for their help with data collection. We also thank Jason
Tylianakis for his comments on the manuscript before submission. JBL
thanks the University of New England for the funding provided to carry
out this work.

\end{document}