### 2021-11-20 OLD/World_Map_Sampling_Location.Rmd
graphical abstract plots

---
title: ""
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE}
#WORLDPLOT WITH LIFE FORM

#Load Libraries
library(readxl)
library(ggplot2)
library(dbplyr)
library(scatterpie)
library(reshape2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

# READ LONG FORMAT DATA 
world <- ne_countries(scale = "medium", returnclass = "sf")

# READ LONG FORMAT DATA

Long_format_metawebs <- readRDS("Long_format_metawebs.RData")

#SUBSET UNIQUE CASES PER NETWORK
Long_format_subset <-  unique(Long_format_metawebs[,c("Plant_species","Id")])
#It seems to work fine
#and the total number of species makes sense, a bit more than the total unique cases for all the networks

#READ TRAIT DATA IN ORDER TO MERGE
data <- read.csv("../Data/Data_processing/Pollinator_species_names/all.csv", stringsAsFactors = F)
data <- data[,c(1,4,6,7,8,51,27)]
levels(as.factor(data$Breeding_system))
#Filter data for columns of interest
data_filtered <- subset(data, Info_level=="flower"|Info_level=="capitulum"|Info_level=="inflorescence"|Info_level=="NA")
#Rename column to merge
colnames(data_filtered)[1] <- "Plant_species"
#Convert to factor 
data_filtered <- as.data.frame(data_filtered, stringsAsFactors = TRUE)

#Merge the two dataframes (trait data and the unique cases per network)
all_species_per_network <- merge(Long_format_subset, data_filtered,  by=c("Plant_species","Id"), all = T )

#Group breeding system per metaweb or web
all_summary_life_form <- dcast(all_species_per_network,Id~life_form,value.var = "life_form")
#Cleaning NA and one unfinished network
all_summary_life_form <- all_summary_life_form[-c(6,30),-c(6)]
#Removing ".csv" from the Id to megere with coordinates
all_summary_life_form$Id <- gsub("\\..*","",all_summary_life_form$Id)

#Now merge with metadata to add coordinates to plot it
metadata <- read.csv("../Data/Data_processing/Metadata/metadata.csv", stringsAsFactors = F)
str(metadata)
metadata_sub <- metadata[,c(1:6,11)]
metadata_sub <- metadata_sub[ , -c(1)]
metadata_sub <- metadata_sub[order(metadata_sub$Id_number),] 
#Fix some coordinates manually, this locations have more than one
#fore representation purposes I will use just one
metadata_sub[6,3] <- "1.575532"
metadata_sub[6,4] <- "52.762395"
metadata_sub[7,3] <- "9.1"
metadata_sub[7,4] <- "56.1"
metadata_sub[29,3] <- "-6.16895"
metadata_sub[29,4] <- "37.234966"
metadata_sub <- metadata_sub[-14,]
colnames(metadata_sub)[1] <- "Id"

str(metadata_sub)

metadata_sub$longitude <- as.numeric(metadata_sub$longitude)
metadata_sub$latitude <- as.numeric(metadata_sub$latitude)

ggplot(data = world) + 
  geom_sf(color = "#E0E0E0", fill = "#EEEEEE", size=0.1,lwd = 0)+
  theme(axis.line = element_line(color='aliceblue'),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())+
  geom_point(data=metadata_sub, aes(longitude,latitude), shape = 16, colour = "#74c365",size = 2.35, alpha=0.9)+
  theme(panel.grid.major = element_line(color = gray(.5), 
  linetype = "dashed", size = 0), panel.background = element_rect(fill = "aliceblue"))+
  xlab("")+ylab("")+
  coord_sf(expand = FALSE)+
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        panel.border = element_blank()
  )

```