### 2023-02-13 Scripts/Pollinator_guild_PCA/Plot_all.Rmd
Trait space with pollinator guilds on it (qualitative version)

---
output: pdf_document
classoption: landscape

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, out.width="85%",  fig.height=12, fig.width=18,fig.align='center'}

#Create panel with all plots, qualitative version

library(ggplot2)

#Load plots
bees = readRDS("../../Data/RData/Bee_PCA.rds")
coleoptera = readRDS("../../Data/RData/Coleoptera_PCA.rds")
lepidoptera = readRDS("../../Data/RData/Lepidoptera_PCA.rds")
non_bee_hymenoptera = readRDS("../../Data/RData/Non_bee_hymenoptera_PCA.rds")
non_syrphid_diptera = readRDS("../../Data/RData/Non_syrphid_diptera_PCA.rds")
syrphids = readRDS("../../Data/RData/Syrphids_PCA.rds")

#Make plot
library(patchwork)

bees + coleoptera + lepidoptera + 
non_bee_hymenoptera + non_syrphid_diptera + syrphids

```