### 2023-02-14 Supplementary_Material/Figures/Anova_Qualitative_Traits.Rmd

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=10, fig.width=10, fig.align='center'}

library(dplyr)
library(kableExtra)

setwd("~/R_Projects/Reproductive Traits")
model1 <- readRDS("Data/RData/model1_tukey_categories.rds")
model2 <- readRDS("Data/RData/model2_tukey_categories.rds")

library(dplyr)
library(kableExtra)
#CREATE TABLE WITH OUTPUT OF ANOVA MODEL

#PC1
am1 <- anova(model1) # save anova output

am1 <- as.data.frame(am1) #convert to dataframe

am2 <- am1[c(1:7),c(2,4,5)] #select columns of interest

am3 <- am2 %>% mutate(across(is.numeric, ~ round(., 2))) #select just two decimals

rownames(am3) <- c("Breeding system","Compatibility system", "Lifespan", "Life form", "Flower shape", "Flower symmetry", "Nectar provision") #rownames

am4 <- tibble::rownames_to_column(am3, "Functional traits")

am4$PC <- c(rep("PC1",7))

#NOW PC2
#PC1
an1 <- anova(model1) # save anova output

an1 <- as.data.frame(an1) #convert to dataframe

an2 <- an1[c(1:7),c(2,4,5)] #select columns of interest

an3 <- an2 %>% mutate(across(is.numeric, ~ round(., 2))) #select just two decimals

rownames(an3) <- c("Breeding system","Compatibility system", "Lifespan", "Life form", "Flower shape", "Flower symmetry", "Nectar provision") #rownames

an4 <- tibble::rownames_to_column(an3, "Functional traits")

an4$PC <- c(rep("PC2",7))

#Merge

d <- rbind(am4,an4)

saveRDS(d, "Data/RData/table_categorical_tukey_test.rds")

#Plot table
#Add latex format in markdown

kable(d, longtable = T,"latex" , booktabs = T,linesep = "\\addlinespace", escape=FALSE) %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

```