### 2023-02-14 Supplementary_Material/Figures/Missing_Data_2_Subset_Floral_Rewards.Rmd

---
output:
  pdf_document: default
  html_document: default

---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE,dpi=2000}

#Set working directory
setwd("~/R_Projects/Reproductive Traits")

########################################################################################################################################################
#SCRIPT FOR DATA IMPUTATION FOR THE SUBSET OF QUANTITATIVE DATA OF NECTAR
########################################################################################################################################################
#1) READ TRAIT DATA

#2) CLEAN DATA AND RENAME LEVELS

#3) EXPLORE PATTERNS OF MISSING DATA

#4) CALCULATE PHYLOGENETIC DISTANCE TO CORRECT WITH EIGENVALUES IN THE IMPUTATION

#5) ICLUDE EIGENVALUES IN RAWDATA TO IMPROVE IMPUTATION OUTPUT

#6) IMPUTE DATA

#7) SAVE DATA
########################################################################################################################################################
#LOAD LIBRARIES
library(readxl) #read excel file (trait data)
library(dplyr) #data manipulation
library(missMDA) #For missing values
library(FactoMineR) #Produce pca biplot with individuals and variables
library(factoextra)
library(ggpubr) 
library(missForest) #random forst imputation
library(rtrees) #for phylogenetic distancelibrary(MASS)
library(ape)
library(PVR) #calculate eigen vectors and add them as a column in the imputation, seems to improve output
library(visdat) #VISUALIZE MISSING DATA
library(tidyverse)
########################################################################################################################################################
#1) READ TRAIT DATA
########################################################################################################################################################
#load data
trait_data <- read_excel("Data/Trait_data_raw/Trait_data_final.xlsx",na = "NA")
########################################################################################################################################################
#2) A) CLEAN DATA AND B) RENAME LEVELS FOR IMPUTATION 
########################################################################################################################################################
#####
#A) CLEAN DATA
#####
#select just filled rows
trait_data_1 <- trait_data[1:1712,]

#filter data, select species with flower level info and capitulum
trait_filtered <- filter(trait_data_1, Info_level == "flower" |  Info_level == "capitulum")
levels(as.factor(trait_filtered$Info_level)) #checking levels
#remove NA's and duplicated species
trait_filtered$Species_all[trait_filtered$Species_all=="NA"]<-NA
trait_filtered_1 <- trait_filtered[!is.na(trait_filtered$Species_all),]
trait_filtered_2 <- trait_filtered_1[!duplicated(trait_filtered_1$Species_all),]

str(trait_filtered_2)
#select columns of interest
t <- trait_filtered_2[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system","Autonomous_selfing_level",
                        "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant",  "Floral_unit_width",
                        "Corolla_diameter_mean", "Corolla_length_mean", "Style_length", "Ovule_number", "life_form", "lifespan",
                        "Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg","Nectar_concentration","Pollen_per_flower")]

########################################################################################################################################################
#2 DATA PREPARATION
########################################################################################################################################################
#RENAME LEVELS 
#Breeding_system
t$Breeding_system <- as.character(t$Breeding_system)
t$Breeding_system[t$Breeding_system=="androdioecious"] <- "dioecious"
t$Breeding_system[t$Breeding_system=="androgynodioecious"] <- "dioecious"
t$Breeding_system[t$Breeding_system=="andromonoecious"] <- "monoecious"
t$Breeding_system[t$Breeding_system=="gynodioecious"] <- "dioecious"
t$Breeding_system[t$Breeding_system=="gynoecious"] <- "monoecious"
t$Breeding_system[t$Breeding_system=="gynomonoecious"] <- "monoecious"
t$Breeding_system[t$Breeding_system=="gynomonoecious"] <- "monoecious"
t$Breeding_system[t$Breeding_system=="monoecious_dioecious"] <- "monoecious"
t$Breeding_system[t$Breeding_system=="polygamo-dioecious"] <- "dioecious"
t$Breeding_system[t$Breeding_system=="protandrous"] <- "hermaphrodite"
t$Breeding_system[t$Breeding_system=="protogynous"] <- "hermaphrodite"
t$Breeding_system[t$Breeding_system=="subdioecious"] <- "dioecious"
t$Breeding_system[t$Breeding_system=="submonoecious"] <- "monoecious"
#change to capital letters
t$Breeding_system[t$Breeding_system=="monoecious"] <- "Monoecious"
t$Breeding_system[t$Breeding_system=="dioecious"] <- "Dioecious"
t$Breeding_system[t$Breeding_system=="hermaphrodite"] <- "Hermaphrodite"
levels(as.factor(t$Breeding_system))
t$Breeding_system <- as.factor(t$Breeding_system)

#Now 3 levels
#Flower_morphology
t$Flower_morphology <- as.character(t$Flower_morphology)
t$Flower_morphology[t$Flower_morphology=="bowl"] <- "open"
t$Flower_morphology[t$Flower_morphology=="dish"] <- "open"
t$Flower_morphology[t$Flower_morphology=="exposed"] <- "open"
t$Flower_morphology[t$Flower_morphology=="spadix"] <- "spike"
t$Flower_morphology[t$Flower_morphology=="open"] <- "Open"
t$Flower_morphology[t$Flower_morphology=="brush"] <- "Brush"
t$Flower_morphology[t$Flower_morphology=="campanulate"] <- "Campanulate"
t$Flower_morphology[t$Flower_morphology=="capitulum"] <- "Capitulum"
t$Flower_morphology[t$Flower_morphology=="funnelform"] <- "Funnelform"
t$Flower_morphology[t$Flower_morphology=="papilionaceous"] <- "Papilionaceous"
t$Flower_morphology[t$Flower_morphology=="spike"] <- "Spike"
t$Flower_morphology[t$Flower_morphology=="tube"] <- "Tube"
levels(as.factor(t$Flower_morphology))
t$Flower_morphology <- as.factor(t$Flower_morphology)

#Life span
t$lifespan <- as.character(t$lifespan )
t$lifespan[t$lifespan=="annual"] <- "short_lived"
t$lifespan[t$lifespan=="biennial"] <- "short_lived"
t$lifespan[t$lifespan=="short_lived"] <- "Short lived"
t$lifespan[t$lifespan=="perennial"] <- "Perennial"
levels(as.factor(t$lifespan))
t$lifespan <- as.factor(t$lifespan )

#use ifelse base r does not work with these NA'S Even with work around. base r does not like the na option of read excel
t$Autonomous_selfing_level_fruit_set <- ifelse(t$Autonomous_selfing_level %in% c("high") & is.na(t$Autonomous_selfing_level_fruit_set), 88, t$Autonomous_selfing_level_fruit_set)
t$Autonomous_selfing_level_fruit_set <- ifelse(t$Autonomous_selfing_level %in% c("medium") & is.na(t$Autonomous_selfing_level_fruit_set), 50.5, t$Autonomous_selfing_level_fruit_set)
t$Autonomous_selfing_level_fruit_set <- ifelse(t$Autonomous_selfing_level %in% c("low") & is.na(t$Autonomous_selfing_level_fruit_set), 13, t$Autonomous_selfing_level_fruit_set)
t$Autonomous_selfing_level_fruit_set <- ifelse(t$Autonomous_selfing_level %in% c("none") & is.na(t$Autonomous_selfing_level_fruit_set), 0, t$Autonomous_selfing_level_fruit_set)

#Now we just select presence/absence of nectar
#select columns of interest
t <- t[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system","Autonomous_selfing_level",
         "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Floral_unit_width",
         "Corolla_diameter_mean", "Corolla_length_mean", "Style_length", "Ovule_number", "life_form", "lifespan",
         "Plant_height_mean_m","Nectar_presence_absence", "Nectar_ul","Nectar_mg","Nectar_concentration","Pollen_per_flower")]

########################################################################################################################################################
#3) EXPLORE PATTERNS OF MISSING DATA
########################################################################################################################################################
#check missing data now
missing_data <- unlist(lapply(t, function(x) sum(is.na(x))))/nrow(t)*100
sort(missing_data[missing_data >= 0], decreasing=T)

#seems ok to impute, just 36% percent of missing data in this column now.

########################################################################################################################################################
#4) CALCULATE PHYLOGENETIC DISTANCE TO CORRECT WITH EIGENVALUES IN THE IMPUTATION
########################################################################################################################################################
#Convert to factor 
cols <- c("Order_all", "Family_all", "Genus_all", "Species_all", "Compatibility_system", "Autonomous_selfing_level", "Flower_morphology",
          "Flower_symmetry", "life_form", "lifespan","Nectar_presence_absence")
t[cols] <- lapply(t[cols], factor)  ## as.factor() could also be used

#remove not found species that do not run with get tree
cols.num <- c("Family_all","Genus_all","Species_all")
t[cols.num] <- sapply(t[cols.num],as.character)
t$Species_all <- gsub("Species_all_", "", t$Species_all)

t <- t[!t$Species_all == "Pinus luchuensis", ]   # remove gymnosperm species

#remove species that are not until species level
#ALL THE SPECIES WITH SP. ARE DELETD
t$Species_geonet <- gsub("M_PL_.*","",t$Species_geonet) #subsitute partial string for nothing
t$Species_geonet <- gsub(" $","", t$Species_geonet, perl=T) #remove trailing spaces
t$Species_geonet <-  sub('sp.1$', 'sp.', t$Species_geonet)#PREPARE ALL SP TO SP.
t$Species_geonet <- sub('sp.2$', 'sp.', t$Species_geonet)#PREPARE ALL SP TO SP.
t$Species_geonet <- sub('sp$', 'sp.', t$Species_geonet) #PREPARE ALL SP TO SP.
t$Species_geonet <- sub("sp.$", "DELETE", t$Species_geonet) #change all sp. for DELETE
t <- t[- grep("DELETE",  t$Species_geonet),] #remove species with "DELETE"
#CHECK LEVELS
levels(factor(t$Species_all))

#Make these NA's as NA
t$Species_all[t$Species_all=="NA"] <- NA
t$Family_all[t$Family_all=="NA"] <- NA
t$Genus_all[t$Genus_all=="NA"] <- NA
#remove NA's
t <- t[!is.na(t$Family_all),]
t <- t[!is.na(t$Species_all),]
t <- t[!is.na(t$Genus_all),]
#Explore missing data
#check missing data now
missing_data <- unlist(lapply(t, function(x) sum(is.na(x))))/nrow(t)*100
sort(missing_data[missing_data >= 0], decreasing=T)

#Subset species that have info for at least one of these floral traits
t_nectar <- t[!is.na(t$Nectar_concentration) | !is.na(t$Nectar_ul)| !is.na(t$Nectar_mg)| !is.na(t$Pollen_per_flower),]
nrow(t_nectar)

#Select just columns with traits
t_nectar <- t_nectar[c("Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg","Nectar_concentration",
         "Pollen_per_flower")]
#################################################
#3)SUMMARY OF THE DATA|ALL COLS CONSIDERED!
#################################################
##Number of filled cells
#nrow(t) * 16 - sum(is.na(t))#[1] 20737
##Dimension of the dataset
#nrow(t) * 16#[1] 24096
##Number of missing values
#sum(is.na(t))#[1] 3359
##Percentage of filled cells
#sum(!is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100#[1] 86.87993
##Percentage of missing values
#sum(is.na(t))/(sum(is.na(t))+sum(!is.na(t)))*100#[1] 13.12007
#
#

#Function to specify number of decimals
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))

no_missing <- specify_decimal(sum(!is.na(t_nectar))/(sum(is.na(t_nectar))+sum(!is.na(t_nectar)))*100,2)
missing <- specify_decimal(sum(is.na(t_nectar))/(sum(is.na(t_nectar))+sum(!is.na(t_nectar)))*100,2)

colnames(t_nectar) <- c("Breeding system","Compatibility system",
         "Autonomous selfing qualitative","Autonomous selfing quantitative", "Flower shape", "Flower symmetry", 
         "Flower number per plant", "Inflorescence width", "Flower width", "Flower length","Style length", "Ovule number", 
         "Life form", "Lifespan","Plant height","Nectar qualitative","Nectar per flower (microlitres)", "Nectar per flower (miligrames)", "Nectar concentration (%)", "Pollen grains per flower")

missing.values <- t_nectar %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <-
  (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
  ggplot() +
  geom_bar(aes(x = reorder(key, desc(pct)), 
               y = pct, fill=isna), 
           stat = 'identity', alpha=0.8) +
  scale_x_discrete(limits = levels) + theme_bw()+
  scale_fill_manual(name = "", 
                    values = c('#57a340', '#d24513'), labels = c("Present values", "Missing values")) +
  coord_flip() +
  labs(title = "", x =
         '', y = "")+  scale_y_continuous(labels = function(x) paste0(x*1, "%")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1))

percentage.plot

```