### 2023-02-14 Supplementary_Material/Tables/Phylo_Signal_Table.Rmd

---
title: ""
output:
  pdf_document: default
  html_document: default
header-includes:  \usepackage{soul}

---

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=10, fig.width=10, fig.align='center'}

library(kableExtra)

setwd("~/R_Projects/Reproductive Traits")

#SELFING
selfing <- readRDS("Data/RData/selfing_phylo.rds")
#FLOWER NUMBER
flower_n <- readRDS("Data/RData/flower_n_phylo.rds")
#INFLO WIDTH
inflow_width <- readRDS("Data/RData/inflow_width_phylo.rds")
#FLOWER WIDTH
flower_width <- readRDS("Data/RData/flower_width_phylo.rds")
#FLOWER LENGHT
flower_length <- readRDS("Data/RData/flower_length_phylo.rds")
#STYLE LENGTH
style_n <- readRDS("Data/RData/style_n_phylo.rds")
#OVULE NUMBER
ovule_n <- readRDS("Data/RData/ovule_n_phylo.rds")
#PLANT HEIGHT
plant_height <- readRDS("Data/RData/plant_height_phylo.rds")
#POLLEN
pollen <- readRDS("Data/RData/pollen_phylo.rds")
#Nectar ul
nectar_ul <- readRDS("Data/RData/Nectar_ul_phylo.rds")
#Nectar mg
nectar_mg <- readRDS("Data/RData/Nectar_mg_phylo.rds")
#Nectar concentration
nectar_con <- readRDS("Data/RData/Nectar_concentration_phylo.rds")

#Select two decimals function
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
#LABELS
label <- c("Autonomous selfing", "Flower number", "Inflorescence width" ,"Flower width", "Flower length", "Style length", "Ovule number", "Plant height", "Nectar ($\\mu$l) per flower", "Nectar (mg) per flower", "Nectar concentration ($\\%$)","Pollen grains per flower")
#LAMBDA
lambda <- c(selfing$lambda , flower_n$lambda, inflow_width$lambda, flower_width$lambda, flower_length$lambda, style_n$lambda, ovule_n$lambda, plant_height$lambda, nectar_ul$lambda, nectar_mg$lambda, nectar_con$lambda, pollen$lambda)
#P-VALUE
p_value <- c(selfing$P , flower_n$P, inflow_width$P,flower_width$P, flower_length$P, style_n$P, ovule_n$P, plant_height$P,nectar_ul$P,
            nectar_mg$P,  nectar_con$P, pollen$P)

tabla <- data.frame(label, specify_decimal(lambda,2), specify_decimal(p_value,2))

colnames(tabla) <- c("Functional traits", "Lambda", "P-value")

kable(tabla, "latex", longtable = T, booktabs = T,linesep = "\\addlinespace", escape=FALSE) %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = "center")

```